### 2015-04-28 inst/examples/synchrony-example.Rmd
Example for get_synchrony function.

# Checking `get_community_synchrony`
A quick check to make sure the function is working as expected.

The `get_community_synchrony` function requires several other packages...
```{r load_libs, echo=TRUE}
library(communitySynchrony)
library(plyr)
library(reshape2)
library(synchrony)
```

### Bring in data
I'm going to use the Kansas data as a test first, then move to the Idaho data.
```{r get_data, echo=TRUE}
site <- "Kansas"
spp_list <- c("BOCU","BOHI","SCSC")
num_spp <- length(spp_list)
ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  ks_data <- rbind(ks_data, spp_data)
} #end species looping for raw data
ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row
```

Chengjin identified several quad-years that need to be removed due to dominance of BOCU.
```{r remove_plots, echo=TRUE}
tmp1<-which(ks_data$quad_data=="q25" & (ks_data$year<35 | ks_data$year>62))
tmp2<-which(ks_data$quad_data=="q27")
tmp3<-which(ks_data$quad=="q28")
tmp4<-which(ks_data$quad=="q30")
tmp5<-which(ks_data$quad=="q31" & (ks_data$year<35 | ks_data$year>39))
tmp6<-which(ks_data$quad=="q32" & (ks_data$year<35 | ks_data$year>41))
tmp<-c(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
ks_data<-ks_data[-tmp,]

# exclude the records later than 1968, to keep the same random year effect...
ks_data<-subset(ks_data,year<68)
```

### Run the function
```{r run_it, echo=TRUE}
out <- get_comm_synchrony(ts_data = ks_data)
str(out)
names(out)
```