### 2015-04-29 docs/prelim_analysis.Rmd
updated with table and plot for all sites

---
output:
  pdf_document:
    fig_caption: yes
---

# Preliminary community synchrony analysis

A couple comments from Claire:

1. In the absence of environmental stochasticity, species fluctuations are not necessarily independent. 
For example in a zero sum game with ecologically equivalent species (Hubbell's neutral model), species fluctuate randomly but they are not independent since the sum is constrained. If there are only two species their fluctuations are necessarily negatively correlated, because the increase in one species must be compensated by a decrease in the other. In a non-neutral model where species have stable equilibrium abundance there are also non independent fluctuations due to species interactions. In competitive communities fluctuations would most of the time have negative correlations overall, but sometimes they can be positive overall.

2. I'm not entirely clear here: null with respect to what? Species interactions, demographic stochasticity and environmental stochasticity can all affect synchrony?

3. Could you show the absolute values, either with three bars per plot, or with a black line that is not at zero but at its unperturbed synchrony value? OK, I see the absolute values in table 1, the differences between sites are quite large compared to the effects of demographic and environmental stochasticities - but it should still be visible. Does this suggest that environmental and demographic stochasticities are minor explanatory variables for community synchrony? What is, then? Diversity?

4. In principle, could the signs of the effects of demographic stochasticity and environmental stochasticity be predicted? For demographic stochasticity, see comment by table 1. For environmental stochasticity: could you quantify the synchrony of species responses to the environment from the synchrony of the intercepts?

5. We could compute the synchrony expected under independent fluctuations: 
$$\text{E}(\phi_{r}) = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }.$$
We could expect demographic stochasticity to tend towards this value: the sign of the effect of demographic stochasticity should be $\text{E}(\phi_{r}) - \phi_{r}$, where $\phi_{r}$ is the observed synchrony of per capita growth rates; removal should be the opposite - does that work?

## Calculate expected and observed sychrony
Under the assumption of independent fluctuations in species' per capit growth rates, we can calculate expected synchrony as:

$$\text{E}(\phi_{r}) = \frac{ \sum_{i=1}^{N}{\sigma_{i}^2} } { \left(\sum_{i=1}^{N}{\sigma_{i}} \right)^2 }$$

where $\sigma_{i}$ is the standard deviation of species $i$'s per capita growth rate through time.

```{r load_libs, echo=FALSE, message=FALSE}
library(communitySynchrony)
library(plyr)
library(reshape2)
library(synchrony)
library(ggplot2)
library(xtable)
library(gridExtra)
library(RColorBrewer)
```

```{r get_all_data, echo=FALSE}
site <- "Kansas"
spp_list <- c("BOCU","BOHI","SCSC")
num_spp <- length(spp_list)
ks_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  ks_data <- rbind(ks_data, spp_data)
} #end species looping for raw data
ks_data <- ks_data[2:nrow(ks_data),] #remove first NA row

tmp1<-which(ks_data$quad_data=="q25" & (ks_data$year<35 | ks_data$year>62))
tmp2<-which(ks_data$quad_data=="q27")
tmp3<-which(ks_data$quad=="q28")
tmp4<-which(ks_data$quad=="q30")
tmp5<-which(ks_data$quad=="q31" & (ks_data$year<35 | ks_data$year>39))
tmp6<-which(ks_data$quad=="q32" & (ks_data$year<35 | ks_data$year>41))
tmp<-c(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
ks_data<-ks_data[-tmp,]

# exclude the records later than 1968, to keep the same random year effect...
ks_data<-subset(ks_data,year<68)

site <- "Idaho"
spp_list <- sort(c("PSSP","HECO","POSE","ARTR"))
num_spp <- length(spp_list)
id_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  id_data <- rbind(id_data, spp_data)
} #end species looping for raw data
id_data <- id_data[2:nrow(id_data),] #remove first NA row
id_data <- subset(id_data, species!="ARTR") #take out the shrub

site <- "Montana"
spp_list <- sort(c("BOGR","HECO","PASM","POSE"))
num_spp <- length(spp_list)
mt_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  mt_data <- rbind(mt_data, spp_data)
} #end species looping for raw data
mt_data <- mt_data[2:nrow(mt_data),] #remove first NA row

site <- "NewMexico"
spp_list <- sort(c("BOER","SPFL"))
num_spp <- length(spp_list)
nm_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  nm_data <- rbind(nm_data, spp_data)
} #end species looping for raw data
nm_data <- nm_data[2:nrow(nm_data),] #remove first NA row

site <- "Arizona"
spp_list <- sort(c("BOER","BORO"))
num_spp <- length(spp_list)
az_data <- data.frame(quad=NA, year=NA, totCover=NA, species=NA)
for(dospp in 1:num_spp){ #loop through species to read in data
  spp_now <- spp_list[dospp]
  quad_file <- paste("../data/", site,"/",spp_now,"/quadratCover.csv",sep="")
  spp_data <- read.csv(quad_file)
  spp_data$species <- spp_now
  az_data <- rbind(az_data, spp_data)
} #end species looping for raw data
az_data <- az_data[2:nrow(az_data),] #remove first NA row
```

```{r get_synchs, echo=FALSE, include=TRUE}
out <- get_comm_synchrony(ts_data = ks_data)
obs_synch <- round(as.numeric(out$pgr_synchrony[1]), 2)
exp_synch <- round(out$pgr_expected_synch_ind_flucts, 2)
exp_demo_effect <- exp_synch - obs_synch
print(out$pgr_synchrony)
print(out$pgr_expected_synch_ind_flucts)
```

```{r sign_text, echo=FALSE}
if(sign(exp_demo_effect)==-1){
  sign_text <- "decrease"
  remove_text <- "increase"
}
if(sign(exp_demo_effect)==1){
 sign_text <- "increase"
 remove_text <- "decrease" 
}
```

So, $\text{E}(\phi_{r})$ = `r exp_synch` and $\phi_{r}$ = `r obs_synch`. Which means the expected effect of demographic stochasticity [$\text{E}(\sigma_{d}) = \text{E}(\phi_{r}) - \phi_{r}$] is `r exp_demo_effect`. So, based on the sign of $\text{E}(\sigma_{d})$ we expect demographic stochasticity to `r sign_text` synchrony among species' per capita growth rates. Thus, removing demographic stochasticity should `r remove_text` synchrony in our population model simulations.

Here are the results for both Kansas and Idaho:

```{r get_all_synch, echo=FALSE, include=TRUE}
out_ks <- out
out_id <- get_comm_synchrony(ts_data = id_data)
out_mt <- get_comm_synchrony(ts_data = mt_data)
out_nm <- get_comm_synchrony(ts_data = nm_data)
out_az <- get_comm_synchrony(ts_data = az_data)
```

```{r synchrony_table, echo=FALSE, include=TRUE, results='asis'}
get_table_metrics <- function(output){
  tmp <- output
  exp_tmp <- round(tmp$pgr_expected_synch_ind_flucts, 2)
  obs_tmp <- round(as.numeric(tmp$pgr_synchrony[1]), 2)
  demo_tmp <- exp_tmp - obs_tmp
  return(c(obs_tmp, exp_tmp, demo_tmp))
}

table_data <- as.data.frame(rbind(get_table_metrics(out_ks),
                                  get_table_metrics(out_id),
                                  get_table_metrics(out_mt),
                                  get_table_metrics(out_nm),
                                  get_table_metrics(out_az)))
table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
table_data <- table_data[,c("Site", "V1", "V2", "V3")]
table_data$spp <- c(length(unique(ks_data$species)),
                    length(unique(id_data$species)),
                    length(unique(mt_data$species)),
                    length(unique(nm_data$species)),
                    length(unique(az_data$species)))
colnames(table_data) <- c("Site",
                          "$\\phi_{r}$", 
                          "$\\text{E}(\\phi_{r})$", 
                          "$\\text{E}(\\sigma_{d})$",
                          "Species richness")

synch_cap <- "Observed and expected synchrony among species' per capita growth rates and expected effect of demographic stochasticity on observed synchrony. Species richness values reflect the number of species analyzed from the community, not the actual richness."
print(xtable(table_data, caption = synch_cap),
      caption.placement="top",
      include.rownames = F, 
      sanitize.colnames.function = identity)
```

```{r time_series_fig, echo=FALSE, include=TRUE, fig.height=7, fig.cap="Observed time series of mean percent cover (averaged over quadrats for each year) and per capita growth rates for each species at each site."}
out_ks$percent_cover$site <- "Kansas"
out_id$percent_cover$site <- "Idaho"
out_mt$percent_cover$site <- "Montana"
out_nm$percent_cover$site <- "New Mexico"
out_az$percent_cover$site <- "Arizona"
cover_data <- rbind(out_ks$percent_cover, out_id$percent_cover,
                    out_mt$percent_cover, out_nm$percent_cover, 
                    out_az$percent_cover)

out_ks$growth_rates$site <- "Kansas"
out_id$growth_rates$site <- "Idaho"
out_mt$growth_rates$site <- "Montana"
out_nm$growth_rates$site <- "New Mexico"
out_az$growth_rates$site <- "Arizona"
pgr_data <- rbind(out_ks$growth_rates, out_id$growth_rates,
                  out_mt$growth_rates, out_nm$growth_rates, 
                  out_az$growth_rates)
all_spp <- unique(pgr_data$species)
color_match <- data.frame(species = all_spp,
                          color = brewer.pal(11,"Paired"))
pgr_data <- merge(pgr_data, color_match, by = "species")
cover_data <- merge(cover_data, color_match, by="species")

cover_plot <- ggplot(data=cover_data, aes(x=(year+1900), y=tot_cover, color=color))+
                geom_line()+
                geom_point()+
                facet_grid(site~., scales = "free_y")+
                guides(color=FALSE)+
                ylab("Mean Percent Cover")+
                xlab("Year")+
                scale_color_manual(values = brewer.pal(11,"Paired"))

pgr_plot <- ggplot(data=pgr_data, aes(x=(year+1900), y=pgr, color=color))+
              geom_line()+
              geom_point()+
              facet_grid(site~., scales = "free_y")+
              guides(color=FALSE)+
              ylab("Per Capita Growth Rate")+
              xlab("Year")+
              scale_color_manual(values = brewer.pal(11,"Paired"))
full_plot <- grid.arrange(cover_plot, pgr_plot, ncol=2)
```