### 2015-05-06 inst/examples/ipm_sims_example.Rmd
new example checking IPM output

---
output: html_document
---
# Checking IPM simulation output
```{r read_data}
library(plyr)
library(reshape2)
library(ggplot2)
output_list <- readRDS("../../results/ipm_simulation_lists.RDS")
mlist <- melt(output_list)
```

## Look at simulated time series
```{r time_series}
ggplot(mlist, aes(x=Var1, y=(value*100), color=Var2))+
  geom_line()+
  facet_grid(L1~L2, scale="free")
```
Hmm... looks like something is wrong with a Kansas species. Also need to compare these results to the one's from Idaho-only script in Diversity-Stability repo.

## Run through synchrony library
```{r get_synchs}
library(communitySynchrony)
get_comm_synchrony(ts_data = subset(mlist, L2=="ENVDEMO" & L1=="Arizona"))
```