### 2016-02-10 docs/appendix_mathsupport.Rmd
adds new appendix for mathematical support

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \usepackage{todonotes}
   - \doublespacing
   - \usepackage[document]{ragged2e}
   - \usepackage{color}
   - \usepackage{rotating}
bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in
linkcolor: black
urlcolor: black

---
# Appendix X for TITLE: Mathematical Support

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

Andrew T. Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}

\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}

\renewcommand*{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

\renewcommand{\theequation}{X\arabic{equation}}

Derivation of Expected Synchrony
--------------------------------
Following @Loreau2013 and @DeMazancourt2013a, we define population growth, ignoring observation error, as

\begin{align}
\tilde{r_i}(t) = r_{mi} \left[ 1- \frac{\tilde{N_i}(t)+\sum_{j \neq i} \alpha_{ij}\tilde{N_j}(t)} {K_i} + \sigma_{ei}\mu_{ei}(t) + \frac{\sigma_{di}\mu_{di}(t)}{\sqrt{\tilde{N_i}(t)}} \right]
\end{align}

\noindent where $\tilde{N_i}(t)$ is the biomass of species *i* in year *t*, and $\tilde{r_i}(t)$ is its instantaneous growth rate in year *t* (tildes denote latent quantities that cannot be known due to observation error). 
$r_{mi}$ is species *i*'s intrinsic rate of increase, $K_i$ is its carrying capacity, and $\alpha_{ij}$ is the interspecific competition coefficient representing the effect of species *j* on species *i*. 
Just as in our vital rate statistical model (Appendix X), environmental stochasticity is incorporated as an additive year effect on a species' growth rate.
In the theoretical population model, environmental stochasticity is incorporated as $\sigma_{ei}\mu_{ei}(t)$, where $\sigma_{ei}^2$ is the environmental variance and $\mu_{ei}$ are normal random variables with zero mean and unit variance that are independent through time but may be correlated.
Demographic stochasticity arises from variations in births and deaths among individuals (e.g., same states, different fates), and is included in the model as a first order, normal approximation [@Lande2003; @DeMazancourt2013a].
$\sigma_{di}^2$ is the demographic variance and $\mu_{di}(t)$ are independent normal variables with zero mean and unit variance.

In the general case, **C** is the variance-covariance matrix for population biomasses at steady state.
It solves the Lyapunov equation $\textbf{C}^\infty = \textbf{A} \textbf{C}^\infty \textbf{A}^T + \textbf{B}$, where

\begin{align}
A_{ij} = 
  \begin{cases}
      1-r_{mi} \frac{N_{i}^*}{K_i}, & i=j \\
      -r_{mi} \frac{N_{i}^*}{K_i} \alpha_{ij}, & i \neq j
    \end{cases}
\end{align}

\noindent and

\begin{equation}
B_{ij} = N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})
\end{equation}

\noindent Similarly, **R** is the variance-covariance matrix for population growth rates at steady state

\begin{equation}
R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} \sum_{k,l} \alpha_{ik} \alpha_{jl} C_{kl} + \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}}.
\end{equation}

\noindent Then, following the synchrony metric of @Loreau2008a, the sychrony of population biomasses is

\begin{equation}
\phi_N = \frac{\sum_{i,j}C_{ij}}{\left( \sum_i \sqrt{C_{ii}} \right)^2}
\end{equation}

\noindent and the synchrony of per capita growth rates is

\begin{equation}
\phi_R = \frac{\sum_{i,j}R_{ij}}{\left( \sum_i \sqrt{R_{ii}} \right)^2}.
\end{equation}

References
----------