### 2016-02-18 docs/appendix_temporalRandomEffects.Rmd
adds temporal random effects table appendix

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \usepackage{todonotes}
   - \doublespacing
   - \usepackage[document]{ragged2e}
   - \usepackage{color}
   - \usepackage{rotating}
bibliography: ~/Dropbox/Bibliography/CommunitySynchrony.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in
linkcolor: black
urlcolor: black

---
# Appendix X for TITLE: Correlations of Species' Random Year Effects

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

Andrew T. Tredennick\footnote{Correspondance: atredenn@gmail.com}\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}

\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Experimental Ecology Station, Centre National de la Recherche Scientifique, Moulis, 09200, France}}

\renewcommand*{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

```{r get_randeffs, echo=FALSE, include=TRUE, warning=FALSE, message=FALSE}
####
####  Clean workspace; Load libraries ------------------------------------------
####
rm(list = ls()) # wipe the workspace clean
library(ggplot2)
library(reshape2)
library(plyr)
library(synchrony)
library(communitySynchrony)
library(xtable)

####
####  Read in growth rate matrices ---------------------------------------------
####
pgr_list <- readRDS("../results/ipm_yearly_pgr.RDS")
site_names <- names(pgr_list)
out_df <- data.frame(site=NA, pgr_synch=NA)
for(do_site in site_names){
  tmp_pgrs <- pgr_list[[do_site]]
  tmp_synch <- as.numeric(community.sync(tmp_pgrs)[1])
  tmp_df <- data.frame(site=do_site, pgr_synch=tmp_synch)
  out_df <- rbind(out_df, tmp_df)
}
pgr_synch <- out_df[2:nrow(out_df),]

####
####  Read in regression fits --------------------------------------------------
####
all_files <- list.files("../results/")
param_files <- all_files[grep("param*", all_files)]
num_vitals <- length(param_files)

list_yr_effects <- list()
for(j in 1:num_vitals){
  param_list <- readRDS(paste0("../results/",param_files[j]))
  vital_name <- unlist(strsplit(param_files[j], "_"))[1]

  if(vital_name != "recruit"){

    site_names <- names(param_list)
    site_year_list <- list()
    for(do_site in site_names){
      tmp_site <- param_list[[do_site]]
      num_spp <- length(tmp_site) 
      tmp_yr_mat <- matrix(ncol = num_spp, nrow=nrow(tmp_site[[1]]))
      for(i in 1:num_spp){
        spp_df <- tmp_site[[i]]
        tmp_yr_mat[,i] <- spp_df$Intercept.yr
      } # end species within site loop
      site_year_list[[do_site]] <- tmp_yr_mat
    } # end site within vital rate loop

  } # end NOT recruit loops

  if(vital_name == "recruit"){

    site_names <- names(param_list)
    site_year_list <- list()
    for(do_site in site_names){
      tmp_site <- param_list[[do_site]]
      tmp_yrs <- tmp_site[grep("intcpt.yr", rownames(tmp_site)),"Mean"]
      stringlist <- strsplit(names(tmp_yrs), ",")
      num_spp <- length(unique(sapply(stringlist, "[[", 2)))
      mat_yrs <- t(matrix(tmp_yrs, nrow=num_spp, byrow = TRUE))
      site_year_list[[do_site]] <- mat_yrs
    } # end site within vital rate loop

  } # end YES recruit loops

  list_yr_effects[[vital_name]] <- site_year_list
} # end vital rate loop

####
####  Calculate average correlation of year effects ----------------------------
####
vital_rates <- names(list_yr_effects)
out_df <- data.frame(site=NA, vital_rate=NA, corr=NA)
for(do_vital in vital_rates){
  tmp_vital <- list_yr_effects[[do_vital]]
  site_names <- names(tmp_vital)
  for(do_site in site_names){
    tmp_site <- tmp_vital[[do_site]]
    tmp_cor <- cor(tmp_site)
    avg_cor <- mean(tmp_cor[upper.tri(tmp_cor)])
    tmp_out <- data.frame(site=do_site, vital_rate=do_vital, corr=avg_cor)
    out_df <- rbind(out_df, tmp_out)
  }
}
cor_df <- out_df[2:nrow(out_df),]
cor_cast <- dcast(cor_df, site~vital_rate)
avg_env_resp <- apply(cor_cast[,2:4], MARGIN = 1, FUN = "mean")

####
####  Calculate synchrony of random year effects -------------------------------
####
vital_rates <- names(list_yr_effects)
out_df <- data.frame(site=NA, vital_rate=NA, synch=NA)
for(do_vital in vital_rates){
  tmp_vital <- list_yr_effects[[do_vital]]
  site_names <- names(tmp_vital)
  for(do_site in site_names){
    tmp_site <- tmp_vital[[do_site]]
    site_synch <- as.numeric(community.sync(tmp_site)[1])
    tmp_out <- data.frame(site=do_site, vital_rate=do_vital, synch=site_synch)
    out_df <- rbind(out_df, tmp_out)
  }
}

out_df <- out_df[2:nrow(out_df),]
out_cast <- dcast(out_df, site~vital_rate)
rm(list=setdiff(ls(), c("out_cast", "out_df", "cor_df", 
                        "cor_cast", "pgr_synch")))

####
####  Save output for tables
####
outlist <- list(synch_long=out_df, synch_wide=out_cast,
                corr_long=cor_df, corr_wide=cor_cast,
                pgr_synch=pgr_synch)
```

```{r make_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE, warning=FALSE}
synch_cap <- "Pearson's correlations among species' temporal random effects by site and vital rate."
print(xtable(cor_cast, caption = synch_cap),
      caption.placement="top",
      include.rownames = F, 
      sanitize.colnames.function = identity,
      comment=FALSE)
```