### 2016-03-23 devel/formula.tex
adds main text word count and tex file

% Created by tikzDevice version 0.10.1 on 2016-02-26 15:02:09
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}
\setlength\PreviewBorder{0pt}
\usepackage{amssymb}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 44.49, 34.03) rectangle (124.53,264.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 44.49, 34.03) rectangle (124.53,264.40);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (108.08,212.04) circle (  3.20);

\path[fill=fillColor] ( 84.15, 75.92) circle (  3.20);

\path[fill=fillColor] ( 93.19, 82.20) circle (  3.20);

\path[fill=fillColor] ( 94.94,165.97) circle (  3.20);

\path[fill=fillColor] (114.56,237.18) circle (  3.20);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] ( 85.24,212.04) circle (  3.20);

\path[fill=fillColor] ( 72.87, 75.92) circle (  3.20);

\path[fill=fillColor] ( 69.23, 82.20) circle (  3.20);

\path[fill=fillColor] ( 72.87,165.97) circle (  3.20);

\path[fill=fillColor] ( 88.87,237.18) circle (  3.20);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 69.23, 89.02) --
	( 69.48, 90.92) --
	( 69.72, 92.82) --
	( 69.97, 94.72) --
	( 70.22, 96.62) --
	( 70.47, 98.53) --
	( 70.72,100.43) --
	( 70.97,102.33) --
	( 71.22,104.23) --
	( 71.46,106.13) --
	( 71.71,108.03) --
	( 71.96,109.93) --
	( 72.21,111.83) --
	( 72.46,113.74) --
	( 72.71,115.64) --
	( 72.96,117.54) --
	( 73.21,119.44) --
	( 73.45,121.34) --
	( 73.70,123.24) --
	( 73.95,125.14) --
	( 74.20,127.05) --
	( 74.45,128.95) --
	( 74.70,130.85) --
	( 74.95,132.75) --
	( 75.20,134.65) --
	( 75.44,136.55) --
	( 75.69,138.45) --
	( 75.94,140.35) --
	( 76.19,142.26) --
	( 76.44,144.16) --
	( 76.69,146.06) --
	( 76.94,147.96) --
	( 77.19,149.86) --
	( 77.43,151.76) --
	( 77.68,153.66) --
	( 77.93,155.56) --
	( 78.18,157.47) --
	( 78.43,159.37) --
	( 78.68,161.27) --
	( 78.93,163.17) --
	( 79.17,165.07) --
	( 79.42,166.97) --
	( 79.67,168.87) --
	( 79.92,170.77) --
	( 80.17,172.68) --
	( 80.42,174.58) --
	( 80.67,176.48) --
	( 80.92,178.38) --
	( 81.16,180.28) --
	( 81.41,182.18) --
	( 81.66,184.08) --
	( 81.91,185.98) --
	( 82.16,187.89) --
	( 82.41,189.79) --
	( 82.66,191.69) --
	( 82.91,193.59) --
	( 83.15,195.49) --
	( 83.40,197.39) --
	( 83.65,199.29) --
	( 83.90,201.19) --
	( 84.15,203.10) --
	( 84.40,205.00) --
	( 84.65,206.90) --
	( 84.89,208.80) --
	( 85.14,210.70) --
	( 85.39,212.60) --
	( 85.64,214.50) --
	( 85.89,216.40) --
	( 86.14,218.31) --
	( 86.39,220.21) --
	( 86.64,222.11) --
	( 86.88,224.01) --
	( 87.13,225.91) --
	( 87.38,227.81) --
	( 87.63,229.71) --
	( 87.88,231.61) --
	( 88.13,233.52) --
	( 88.38,235.42) --
	( 88.63,237.32) --
	( 88.87,239.22);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 84.15, 71.19) --
	( 84.53, 73.36) --
	( 84.92, 75.52) --
	( 85.30, 77.69) --
	( 85.69, 79.86) --
	( 86.07, 82.02) --
	( 86.46, 84.19) --
	( 86.84, 86.35) --
	( 87.23, 88.52) --
	( 87.61, 90.68) --
	( 88.00, 92.85) --
	( 88.38, 95.02) --
	( 88.77, 97.18) --
	( 89.15, 99.35) --
	( 89.54,101.51) --
	( 89.92,103.68) --
	( 90.31,105.84) --
	( 90.69,108.01) --
	( 91.08,110.18) --
	( 91.46,112.34) --
	( 91.85,114.51) --
	( 92.23,116.67) --
	( 92.62,118.84) --
	( 93.00,121.00) --
	( 93.39,123.17) --
	( 93.77,125.34) --
	( 94.16,127.50) --
	( 94.54,129.67) --
	( 94.93,131.83) --
	( 95.31,134.00) --
	( 95.70,136.16) --
	( 96.08,138.33) --
	( 96.46,140.50) --
	( 96.85,142.66) --
	( 97.23,144.83) --
	( 97.62,146.99) --
	( 98.00,149.16) --
	( 98.39,151.33) --
	( 98.77,153.49) --
	( 99.16,155.66) --
	( 99.54,157.82) --
	( 99.93,159.99) --
	(100.31,162.15) --
	(100.70,164.32) --
	(101.08,166.49) --
	(101.47,168.65) --
	(101.85,170.82) --
	(102.24,172.98) --
	(102.62,175.15) --
	(103.01,177.31) --
	(103.39,179.48) --
	(103.78,181.65) --
	(104.16,183.81) --
	(104.55,185.98) --
	(104.93,188.14) --
	(105.32,190.31) --
	(105.70,192.47) --
	(106.09,194.64) --
	(106.47,196.81) --
	(106.86,198.97) --
	(107.24,201.14) --
	(107.63,203.30) --
	(108.01,205.47) --
	(108.40,207.63) --
	(108.78,209.80) --
	(109.17,211.97) --
	(109.55,214.13) --
	(109.94,216.30) --
	(110.32,218.46) --
	(110.71,220.63) --
	(111.09,222.80) --
	(111.48,224.96) --
	(111.86,227.13) --
	(112.25,229.29) --
	(112.63,231.46) --
	(113.02,233.62) --
	(113.40,235.79) --
	(113.79,237.96) --
	(114.17,240.12) --
	(114.56,242.29);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 44.49, 34.03) -- (124.53,264.40);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 44.49, 34.03) rectangle (124.53,264.40);
\end{scope}
\begin{scope}
\path[clip] (127.54, 34.03) rectangle (207.59,264.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (127.54, 34.03) rectangle (207.59,264.40);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (190.57,230.89) circle (  3.20);

\path[fill=fillColor] (162.76,157.59) circle (  3.20);

\path[fill=fillColor] (176.06,119.90) circle (  3.20);

\path[fill=fillColor] (165.59,155.50) circle (  3.20);

\path[fill=fillColor] (193.62,224.61) circle (  3.20);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] (168.29,230.89) circle (  3.20);

\path[fill=fillColor] (155.92,157.59) circle (  3.20);

\path[fill=fillColor] (152.28,119.90) circle (  3.20);

\path[fill=fillColor] (155.92,155.50) circle (  3.20);

\path[fill=fillColor] (171.93,224.61) circle (  3.20);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] (152.28,131.37) --
	(152.53,132.71) --
	(152.78,134.05) --
	(153.03,135.40) --
	(153.28,136.74) --
	(153.53,138.08) --
	(153.78,139.42) --
	(154.02,140.76) --
	(154.27,142.11) --
	(154.52,143.45) --
	(154.77,144.79) --
	(155.02,146.13) --
	(155.27,147.47) --
	(155.52,148.81) --
	(155.77,150.16) --
	(156.01,151.50) --
	(156.26,152.84) --
	(156.51,154.18) --
	(156.76,155.52) --
	(157.01,156.87) --
	(157.26,158.21) --
	(157.51,159.55) --
	(157.75,160.89) --
	(158.00,162.23) --
	(158.25,163.58) --
	(158.50,164.92) --
	(158.75,166.26) --
	(159.00,167.60) --
	(159.25,168.94) --
	(159.50,170.28) --
	(159.74,171.63) --
	(159.99,172.97) --
	(160.24,174.31) --
	(160.49,175.65) --
	(160.74,176.99) --
	(160.99,178.34) --
	(161.24,179.68) --
	(161.49,181.02) --
	(161.73,182.36) --
	(161.98,183.70) --
	(162.23,185.04) --
	(162.48,186.39) --
	(162.73,187.73) --
	(162.98,189.07) --
	(163.23,190.41) --
	(163.47,191.75) --
	(163.72,193.10) --
	(163.97,194.44) --
	(164.22,195.78) --
	(164.47,197.12) --
	(164.72,198.46) --
	(164.97,199.80) --
	(165.22,201.15) --
	(165.46,202.49) --
	(165.71,203.83) --
	(165.96,205.17) --
	(166.21,206.51) --
	(166.46,207.86) --
	(166.71,209.20) --
	(166.96,210.54) --
	(167.21,211.88) --
	(167.45,213.22) --
	(167.70,214.57) --
	(167.95,215.91) --
	(168.20,217.25) --
	(168.45,218.59) --
	(168.70,219.93) --
	(168.95,221.27) --
	(169.19,222.62) --
	(169.44,223.96) --
	(169.69,225.30) --
	(169.94,226.64) --
	(170.19,227.98) --
	(170.44,229.33) --
	(170.69,230.67) --
	(170.94,232.01) --
	(171.18,233.35) --
	(171.43,234.69) --
	(171.68,236.03) --
	(171.93,237.38);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] (162.76,138.10) --
	(163.15,139.13) --
	(163.54,140.17) --
	(163.94,141.20) --
	(164.33,142.23) --
	(164.72,143.27) --
	(165.11,144.30) --
	(165.50,145.34) --
	(165.89,146.37) --
	(166.28,147.40) --
	(166.67,148.44) --
	(167.06,149.47) --
	(167.45,150.51) --
	(167.84,151.54) --
	(168.23,152.57) --
	(168.62,153.61) --
	(169.01,154.64) --
	(169.40,155.68) --
	(169.79,156.71) --
	(170.18,157.74) --
	(170.57,158.78) --
	(170.96,159.81) --
	(171.36,160.85) --
	(171.75,161.88) --
	(172.14,162.91) --
	(172.53,163.95) --
	(172.92,164.98) --
	(173.31,166.02) --
	(173.70,167.05) --
	(174.09,168.08) --
	(174.48,169.12) --
	(174.87,170.15) --
	(175.26,171.19) --
	(175.65,172.22) --
	(176.04,173.25) --
	(176.43,174.29) --
	(176.82,175.32) --
	(177.21,176.36) --
	(177.60,177.39) --
	(177.99,178.42) --
	(178.38,179.46) --
	(178.78,180.49) --
	(179.17,181.53) --
	(179.56,182.56) --
	(179.95,183.59) --
	(180.34,184.63) --
	(180.73,185.66) --
	(181.12,186.70) --
	(181.51,187.73) --
	(181.90,188.76) --
	(182.29,189.80) --
	(182.68,190.83) --
	(183.07,191.87) --
	(183.46,192.90) --
	(183.85,193.93) --
	(184.24,194.97) --
	(184.63,196.00) --
	(185.02,197.04) --
	(185.41,198.07) --
	(185.80,199.10) --
	(186.20,200.14) --
	(186.59,201.17) --
	(186.98,202.21) --
	(187.37,203.24) --
	(187.76,204.27) --
	(188.15,205.31) --
	(188.54,206.34) --
	(188.93,207.38) --
	(189.32,208.41) --
	(189.71,209.44) --
	(190.10,210.48) --
	(190.49,211.51) --
	(190.88,212.55) --
	(191.27,213.58) --
	(191.66,214.61) --
	(192.05,215.65) --
	(192.44,216.68) --
	(192.83,217.72) --
	(193.23,218.75) --
	(193.62,219.78);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (127.54, 34.03) -- (207.59,264.40);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (127.54, 34.03) rectangle (207.59,264.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 44.49,264.40) rectangle (124.53,277.04);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 84.51,267.41) {cover};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (127.54,264.40) rectangle (207.59,277.04);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (167.56,267.41) {growth rate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37, 41.20) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37, 93.56) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,145.91) {0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,198.27) {0.75};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,250.62) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22, 44.51) --
	( 44.49, 44.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22, 96.86) --
	( 44.49, 96.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,149.22) --
	( 44.49,149.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,201.57) --
	( 44.49,201.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,253.93) --
	( 44.49,253.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.12, 29.77) --
	( 48.12, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 66.32, 29.77) --
	( 66.32, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.51, 29.77) --
	( 84.51, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.70, 29.77) --
	(102.70, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.89, 29.77) --
	(120.89, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 48.12, 20.31) {0.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 66.32, 20.31) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 84.51, 20.31) {0.50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (102.70, 20.31) {0.75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (120.89, 20.31) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] (131.18, 29.77) --
	(131.18, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.37, 29.77) --
	(149.37, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.56, 29.77) --
	(167.56, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (185.76, 29.77) --
	(185.76, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.95, 29.77) --
	(203.95, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (131.18, 20.31) {0.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (149.37, 20.31) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (167.56, 20.31) {0.50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (185.76, 20.31) {0.75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (203.95, 20.31) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (126.04,  9.03) {Predicted Synchrony};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 17.30,149.22) {Observed Synchrony};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (216.45,125.38) rectangle (268.17,173.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (220.72,162.17) {\bfseries predtype};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.72,144.10) rectangle (235.18,158.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] (227.95,151.33) circle (  3.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] (222.17,151.33) -- (233.73,151.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.72,129.64) rectangle (235.18,144.10);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (227.95,136.87) circle (  3.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] (222.17,136.87) -- (233.73,136.87);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (236.98,148.02) {$\mathcal{M}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (236.98,133.57) {NA};
\end{scope}
\end{tikzpicture}
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 44.49, 34.03) rectangle (124.53,264.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 44.49, 34.03) rectangle (124.53,264.40);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (108.08,212.04) circle (  3.20);

\path[fill=fillColor] ( 84.15, 75.92) circle (  3.20);

\path[fill=fillColor] ( 93.19, 82.20) circle (  3.20);

\path[fill=fillColor] ( 94.94,165.97) circle (  3.20);

\path[fill=fillColor] (114.56,237.18) circle (  3.20);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] ( 85.24,212.04) circle (  3.20);

\path[fill=fillColor] ( 72.87, 75.92) circle (  3.20);

\path[fill=fillColor] ( 69.23, 82.20) circle (  3.20);

\path[fill=fillColor] ( 72.87,165.97) circle (  3.20);

\path[fill=fillColor] ( 88.87,237.18) circle (  3.20);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 69.23, 89.02) --
	( 69.48, 90.92) --
	( 69.72, 92.82) --
	( 69.97, 94.72) --
	( 70.22, 96.62) --
	( 70.47, 98.53) --
	( 70.72,100.43) --
	( 70.97,102.33) --
	( 71.22,104.23) --
	( 71.46,106.13) --
	( 71.71,108.03) --
	( 71.96,109.93) --
	( 72.21,111.83) --
	( 72.46,113.74) --
	( 72.71,115.64) --
	( 72.96,117.54) --
	( 73.21,119.44) --
	( 73.45,121.34) --
	( 73.70,123.24) --
	( 73.95,125.14) --
	( 74.20,127.05) --
	( 74.45,128.95) --
	( 74.70,130.85) --
	( 74.95,132.75) --
	( 75.20,134.65) --
	( 75.44,136.55) --
	( 75.69,138.45) --
	( 75.94,140.35) --
	( 76.19,142.26) --
	( 76.44,144.16) --
	( 76.69,146.06) --
	( 76.94,147.96) --
	( 77.19,149.86) --
	( 77.43,151.76) --
	( 77.68,153.66) --
	( 77.93,155.56) --
	( 78.18,157.47) --
	( 78.43,159.37) --
	( 78.68,161.27) --
	( 78.93,163.17) --
	( 79.17,165.07) --
	( 79.42,166.97) --
	( 79.67,168.87) --
	( 79.92,170.77) --
	( 80.17,172.68) --
	( 80.42,174.58) --
	( 80.67,176.48) --
	( 80.92,178.38) --
	( 81.16,180.28) --
	( 81.41,182.18) --
	( 81.66,184.08) --
	( 81.91,185.98) --
	( 82.16,187.89) --
	( 82.41,189.79) --
	( 82.66,191.69) --
	( 82.91,193.59) --
	( 83.15,195.49) --
	( 83.40,197.39) --
	( 83.65,199.29) --
	( 83.90,201.19) --
	( 84.15,203.10) --
	( 84.40,205.00) --
	( 84.65,206.90) --
	( 84.89,208.80) --
	( 85.14,210.70) --
	( 85.39,212.60) --
	( 85.64,214.50) --
	( 85.89,216.40) --
	( 86.14,218.31) --
	( 86.39,220.21) --
	( 86.64,222.11) --
	( 86.88,224.01) --
	( 87.13,225.91) --
	( 87.38,227.81) --
	( 87.63,229.71) --
	( 87.88,231.61) --
	( 88.13,233.52) --
	( 88.38,235.42) --
	( 88.63,237.32) --
	( 88.87,239.22);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 84.15, 71.19) --
	( 84.53, 73.36) --
	( 84.92, 75.52) --
	( 85.30, 77.69) --
	( 85.69, 79.86) --
	( 86.07, 82.02) --
	( 86.46, 84.19) --
	( 86.84, 86.35) --
	( 87.23, 88.52) --
	( 87.61, 90.68) --
	( 88.00, 92.85) --
	( 88.38, 95.02) --
	( 88.77, 97.18) --
	( 89.15, 99.35) --
	( 89.54,101.51) --
	( 89.92,103.68) --
	( 90.31,105.84) --
	( 90.69,108.01) --
	( 91.08,110.18) --
	( 91.46,112.34) --
	( 91.85,114.51) --
	( 92.23,116.67) --
	( 92.62,118.84) --
	( 93.00,121.00) --
	( 93.39,123.17) --
	( 93.77,125.34) --
	( 94.16,127.50) --
	( 94.54,129.67) --
	( 94.93,131.83) --
	( 95.31,134.00) --
	( 95.70,136.16) --
	( 96.08,138.33) --
	( 96.46,140.50) --
	( 96.85,142.66) --
	( 97.23,144.83) --
	( 97.62,146.99) --
	( 98.00,149.16) --
	( 98.39,151.33) --
	( 98.77,153.49) --
	( 99.16,155.66) --
	( 99.54,157.82) --
	( 99.93,159.99) --
	(100.31,162.15) --
	(100.70,164.32) --
	(101.08,166.49) --
	(101.47,168.65) --
	(101.85,170.82) --
	(102.24,172.98) --
	(102.62,175.15) --
	(103.01,177.31) --
	(103.39,179.48) --
	(103.78,181.65) --
	(104.16,183.81) --
	(104.55,185.98) --
	(104.93,188.14) --
	(105.32,190.31) --
	(105.70,192.47) --
	(106.09,194.64) --
	(106.47,196.81) --
	(106.86,198.97) --
	(107.24,201.14) --
	(107.63,203.30) --
	(108.01,205.47) --
	(108.40,207.63) --
	(108.78,209.80) --
	(109.17,211.97) --
	(109.55,214.13) --
	(109.94,216.30) --
	(110.32,218.46) --
	(110.71,220.63) --
	(111.09,222.80) --
	(111.48,224.96) --
	(111.86,227.13) --
	(112.25,229.29) --
	(112.63,231.46) --
	(113.02,233.62) --
	(113.40,235.79) --
	(113.79,237.96) --
	(114.17,240.12) --
	(114.56,242.29);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 44.49, 34.03) -- (124.53,264.40);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 44.49, 34.03) rectangle (124.53,264.40);
\end{scope}
\begin{scope}
\path[clip] (127.54, 34.03) rectangle (207.59,264.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (127.54, 34.03) rectangle (207.59,264.40);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (190.57,230.89) circle (  3.20);

\path[fill=fillColor] (162.76,157.59) circle (  3.20);

\path[fill=fillColor] (176.06,119.90) circle (  3.20);

\path[fill=fillColor] (165.59,155.50) circle (  3.20);

\path[fill=fillColor] (193.62,224.61) circle (  3.20);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] (168.29,230.89) circle (  3.20);

\path[fill=fillColor] (155.92,157.59) circle (  3.20);

\path[fill=fillColor] (152.28,119.90) circle (  3.20);

\path[fill=fillColor] (155.92,155.50) circle (  3.20);

\path[fill=fillColor] (171.93,224.61) circle (  3.20);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] (152.28,131.37) --
	(152.53,132.71) --
	(152.78,134.05) --
	(153.03,135.40) --
	(153.28,136.74) --
	(153.53,138.08) --
	(153.78,139.42) --
	(154.02,140.76) --
	(154.27,142.11) --
	(154.52,143.45) --
	(154.77,144.79) --
	(155.02,146.13) --
	(155.27,147.47) --
	(155.52,148.81) --
	(155.77,150.16) --
	(156.01,151.50) --
	(156.26,152.84) --
	(156.51,154.18) --
	(156.76,155.52) --
	(157.01,156.87) --
	(157.26,158.21) --
	(157.51,159.55) --
	(157.75,160.89) --
	(158.00,162.23) --
	(158.25,163.58) --
	(158.50,164.92) --
	(158.75,166.26) --
	(159.00,167.60) --
	(159.25,168.94) --
	(159.50,170.28) --
	(159.74,171.63) --
	(159.99,172.97) --
	(160.24,174.31) --
	(160.49,175.65) --
	(160.74,176.99) --
	(160.99,178.34) --
	(161.24,179.68) --
	(161.49,181.02) --
	(161.73,182.36) --
	(161.98,183.70) --
	(162.23,185.04) --
	(162.48,186.39) --
	(162.73,187.73) --
	(162.98,189.07) --
	(163.23,190.41) --
	(163.47,191.75) --
	(163.72,193.10) --
	(163.97,194.44) --
	(164.22,195.78) --
	(164.47,197.12) --
	(164.72,198.46) --
	(164.97,199.80) --
	(165.22,201.15) --
	(165.46,202.49) --
	(165.71,203.83) --
	(165.96,205.17) --
	(166.21,206.51) --
	(166.46,207.86) --
	(166.71,209.20) --
	(166.96,210.54) --
	(167.21,211.88) --
	(167.45,213.22) --
	(167.70,214.57) --
	(167.95,215.91) --
	(168.20,217.25) --
	(168.45,218.59) --
	(168.70,219.93) --
	(168.95,221.27) --
	(169.19,222.62) --
	(169.44,223.96) --
	(169.69,225.30) --
	(169.94,226.64) --
	(170.19,227.98) --
	(170.44,229.33) --
	(170.69,230.67) --
	(170.94,232.01) --
	(171.18,233.35) --
	(171.43,234.69) --
	(171.68,236.03) --
	(171.93,237.38);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] (162.76,138.10) --
	(163.15,139.13) --
	(163.54,140.17) --
	(163.94,141.20) --
	(164.33,142.23) --
	(164.72,143.27) --
	(165.11,144.30) --
	(165.50,145.34) --
	(165.89,146.37) --
	(166.28,147.40) --
	(166.67,148.44) --
	(167.06,149.47) --
	(167.45,150.51) --
	(167.84,151.54) --
	(168.23,152.57) --
	(168.62,153.61) --
	(169.01,154.64) --
	(169.40,155.68) --
	(169.79,156.71) --
	(170.18,157.74) --
	(170.57,158.78) --
	(170.96,159.81) --
	(171.36,160.85) --
	(171.75,161.88) --
	(172.14,162.91) --
	(172.53,163.95) --
	(172.92,164.98) --
	(173.31,166.02) --
	(173.70,167.05) --
	(174.09,168.08) --
	(174.48,169.12) --
	(174.87,170.15) --
	(175.26,171.19) --
	(175.65,172.22) --
	(176.04,173.25) --
	(176.43,174.29) --
	(176.82,175.32) --
	(177.21,176.36) --
	(177.60,177.39) --
	(177.99,178.42) --
	(178.38,179.46) --
	(178.78,180.49) --
	(179.17,181.53) --
	(179.56,182.56) --
	(179.95,183.59) --
	(180.34,184.63) --
	(180.73,185.66) --
	(181.12,186.70) --
	(181.51,187.73) --
	(181.90,188.76) --
	(182.29,189.80) --
	(182.68,190.83) --
	(183.07,191.87) --
	(183.46,192.90) --
	(183.85,193.93) --
	(184.24,194.97) --
	(184.63,196.00) --
	(185.02,197.04) --
	(185.41,198.07) --
	(185.80,199.10) --
	(186.20,200.14) --
	(186.59,201.17) --
	(186.98,202.21) --
	(187.37,203.24) --
	(187.76,204.27) --
	(188.15,205.31) --
	(188.54,206.34) --
	(188.93,207.38) --
	(189.32,208.41) --
	(189.71,209.44) --
	(190.10,210.48) --
	(190.49,211.51) --
	(190.88,212.55) --
	(191.27,213.58) --
	(191.66,214.61) --
	(192.05,215.65) --
	(192.44,216.68) --
	(192.83,217.72) --
	(193.23,218.75) --
	(193.62,219.78);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (127.54, 34.03) -- (207.59,264.40);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (127.54, 34.03) rectangle (207.59,264.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 44.49,264.40) rectangle (124.53,277.04);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 84.51,267.41) {cover};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (127.54,264.40) rectangle (207.59,277.04);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (167.56,267.41) {growth rate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37, 41.20) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37, 93.56) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,145.91) {0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,198.27) {0.75};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 37.37,250.62) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22, 44.51) --
	( 44.49, 44.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22, 96.86) --
	( 44.49, 96.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,149.22) --
	( 44.49,149.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,201.57) --
	( 44.49,201.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.22,253.93) --
	( 44.49,253.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.12, 29.77) --
	( 48.12, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 66.32, 29.77) --
	( 66.32, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.51, 29.77) --
	( 84.51, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.70, 29.77) --
	(102.70, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.89, 29.77) --
	(120.89, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 48.12, 20.31) {0.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 66.32, 20.31) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 84.51, 20.31) {0.50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (102.70, 20.31) {0.75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (120.89, 20.31) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{gray}{0.45}

\path[draw=drawColor,line width= 0.6pt,line join=round] (131.18, 29.77) --
	(131.18, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.37, 29.77) --
	(149.37, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.56, 29.77) --
	(167.56, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (185.76, 29.77) --
	(185.76, 34.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.95, 29.77) --
	(203.95, 34.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (131.18, 20.31) {0.00};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (149.37, 20.31) {0.25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (167.56, 20.31) {0.50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (185.76, 20.31) {0.75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (203.95, 20.31) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (126.04,  9.03) {Predicted Synchrony};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 17.30,149.22) {Observed Synchrony};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (216.45,125.38) rectangle (268.17,173.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (220.72,162.17) {\bfseries predtype};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.72,144.10) rectangle (235.18,158.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{157,97,136}

\path[fill=fillColor] (227.95,151.33) circle (  3.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{157,97,136}

\path[draw=drawColor,line width= 1.1pt,line join=round] (222.17,151.33) -- (233.73,151.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.72,129.64) rectangle (235.18,144.10);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{fillColor}{RGB}{151,168,97}

\path[fill=fillColor] (227.95,136.87) circle (  3.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{151,168,97}

\path[draw=drawColor,line width= 1.1pt,line join=round] (222.17,136.87) -- (233.73,136.87);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{1,2,2}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (236.98,149.14) {M};
\end{scope}
\end{tikzpicture}

\end{document}