### 2016-07-02 docs/ms_communitySynchrony.tex
final edits for ELE resubmission

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{todonotes}
\doublespacing
\usepackage{rotating}
\usepackage{color, soul}

\begin{document}

\maketitle

\begin{singlespace}

\begin{center}
\large{\textbf{Environmental responses, not species interactions, determine species synchrony in natural plant communities}}

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\vspace{1em}

\normalsize{Andrew T. Tredennick\textsuperscript{1}, Claire de Mazancourt\textsuperscript{2}, Michel Loreau\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}

\vspace{1em}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA}}

\textit{\small{\textsuperscript{2}Centre for Biodiversity Theory and Modelling, Theoretical and Experimental Ecology Station, CNRS and Paul Sabatier University, Moulis, 09200, France}}

\end{center}

\vspace{2em}

\textbf{Keywords}: synchrony, compensatory dynamics, environmental stochasticity, demographic stochasticity, interspecific competition, stability

\vspace{2em}

\textbf{Authorship}: ATT, CdM, and PBA designed the research. CdM and ML designed and analyzed the theoretical model. ATT and PBA performed the empirical analysis. All authors contributed to the interpretation and presentation of results. ATT wrote the first draft of the manuscript, and all authors contributed to manuscript revisions.

\textbf{Running Title}: Drivers of species synchrony in natural plant communities

\textbf{Article Type}: Letter

\textbf{Number of Words}: Abstract = 136; Main Text = 4992

\textbf{Number of References}: 46

\textbf{Number of Tables and Figures}:  2 Tables;  3 Figures

\textbf{Corresponding Author}: \\
\hspace{2em}Andrew Tredennick \\  
\hspace{2em}Department of Wildland Resources and the Ecology Center\\
\hspace{2em}Utah State University \\
\hspace{2em}5230 Old Main Hill \\
\hspace{2em}Logan, Utah 84322 USA \\
\hspace{2em}Phone: +1-970-443-1599 \\
\hspace{2em}Fax: +1-435-797-3796 \\
\hspace{2em}Email: atredenn@gmail.com

\end{singlespace}\renewcommand*{\thefootnote}{\arabic{footnote}}

\setcounter{footnote}{0}

\newpage{}

\begin{abstract}
Temporal asynchrony among species \textcolor{blue}{helps} diversity \textcolor{blue}{to} stabilize ecosystem functioning, but identifying the mechanisms that determine synchrony remains a challenge.
Here, we refine and test theory showing that synchrony depends on three factors: species responses to environmental variation, interspecific \textcolor{blue}{interactions}, and demographic stochasticity. We then conduct simulation experiments with empirical population models to quantify the relative importance of these factors in five plant communities.
\textcolor{blue}{We found} that the average synchrony of per capita growth rates, which can range from 0 (perfect asynchrony) to 1 (perfect synchrony), was higher when environmental variation was present (0.62) rather than absent (0.43).
Removing interspecific \textcolor{blue}{interactions} and demographic stochasticity had small effects on synchrony. In these plant communities, where species interactions and demographic stochasticity have little influence, synchrony reflects the covariance in species responses to the environment.
\end{abstract}

\newpage{} \setlength{\parindent}{5ex}

\subsection{INTRODUCTION}\label{introduction}

Ecosystems are being transformed by species extinctions (Cardinale et
al. 2012), changes in community composition (Vellend et al. 2013,
Dornelas et al. 2014), and anthropogenic environmental change (Vitousek
et al. 1997), impacting the provisioning and stability of ecosystem
services (Loreau et al. 2001, Hooper et al. 2005, Rockstrom et al.
2009). Experiments have provided compelling evidence that decreases in
species richness will decrease productivity (Tilman et al. 2001) and the
temporal stability of productivity (Tilman et al. 2006, Hector et al.
2010). The stabilizing effect of species richness stems from individual
species responding \textcolor{blue}{in different ways} to environmental
fluctuations (environmental stochasticity), or fluctuating
asynchronously because of random chance events (demographic
stochasticity) (Isbell et al. 2009, Hector et al. 2010, {{de
Mazancourt}} et al. 2013). Species richness affects synchrony because
larger species pools are more likely to contain species that respond
disimilarly to environmental conditions (Yachi and Loreau 1999),
implying that species losses will reduce ecosystem stability. Even
without species losses, abiotic homogenization can weaken compensatory
dynamics and, in turn, decrease temporal stability of ecosystem
functioning (Hautier et al. 2014).
\textcolor{blue}{The link between synchrony and stability means that a mechanistic understanding of synchrony can help us predict the impacts of global change on ecosystem stability.}

Theory identifies three main determinants of species synchrony:
environmental stochasticity, demographic stochasticity, and
\textcolor{blue}{interspecific interactions} (Loreau and {{de
Mazancourt}} 2008, 2013, Gonzalez and Loreau 2009). For example, in a
community composed of large \textcolor{blue}{populations} (no
demographic stochasticity)
\textcolor{blue}{with weak interspecific interations}, community-wide
species synchrony should be determined by the covariance of species'
responses to the environment (Loreau and {{de Mazancourt}} 2008).
However, \textcolor{blue}{this} prediction relies on a relatively simple
population model and only holds under two assumptions: (i) species'
responses to the environment are similar in magnitude and (ii) all
species have similar growth rates.
\textcolor{blue}{Whether such theoretical predictions hold in natural communities where species differences are unlikely to be symmetrical is unkown because few studies have explicitly tested theory on the drivers of species synchrony in natural communities}
(Mutshinda et al. 2009, Thibaut et al. 2012),
\textcolor{blue}{and they did not consider demographic stochasiticity.}

In grasslands, most empirical studies have focused on whether species
synchrony is primarily an outcome of species-specific responses to
environmental conditions (Hautier et al. 2014) or competition (Gross et
al. 2014). Even beyond grassland studies, whether competition or
environmental responses drive compensatory dynamics remains
controversial (reviewed in Gonzalez and Loreau 2009). In part,
controversy remains because quantifying the relative strengths of each
driver on the degree of synchrony from
\textcolor{blue}{the covariance matrix of species abundances} (e.g.,
Houlahan et al. 2007) is impossible. This is because an unbiased null
expectation for synchrony does not exist (Loreau and {{de Mazancourt}}
2008) and observed synchrony can arise from non-unique combinations of
factors (Ranta et al. 2008). For example, weak synchrony of population
abundances could reflect positive environmental correlations
(synchronizing effect) offset by strong competition (desynchronizing
effect), or negative environmental correlations and weak competition.

The best way to quantify the effects of environmental stochasticity,
demographic stochasticity, and interspecific
\textcolor{blue}{interactions} is to remove them one-by-one, and in
combination. In principle, this could be done in an extremely controlled
laboratory setting, but empirically-based models of interacting
populations, fit with data sets from natural communities, offer a
practical alternative. For example, Mutshinda et al. (2009) fit a
dynamic population model to several community time series of insect and
bird abundances. They used a statistical technique to decompose temporal
variation into competition and environmental components, and found that
positively correlated environmental responses among species determined
community dynamics. Thibaut et al. (2012)
\textcolor{blue}{used a similar approach for reef fish and came to a similar conclusion: environmental responses determine synchrony.}
While a major step forward, Mutshinda et al.'s (2009) and Thibaut et
al.'s (2012) modeling technique relied on abundance data that may or may
not reliably capture competitive interactions that occur at the
individual level. Furthermore, although both studies quantified the
relative importance of environmental stochasticity and
\textcolor{blue}{interspecific interactions} to explain the observed
variation of species synchrony, they did not use the model to quantify
how much synchrony would change when each factor is removed.

Here, we use multi-species population models fit to long-term
demographic data from five semi-arid plant communities to test theory on
the drivers of species synchrony. Our objectives are to (1) derive and
test theoretical predictions of species synchrony and (2) determine the
relative influence of environmental stochasticity, demographic
stochasticity, and interspecific \textcolor{blue}{interactions} on
species synchrony in natural plant communities. To achieve these
objectives, we first refine theory that has been used to predict the
effects of species richness on ecosystem stability ({{de Mazancourt}} et
al. 2013) and species synchrony (Loreau and {{de Mazancourt}} 2008) to
generate predictions of community-wide species synchrony under two
limiting cases derived from the dynamics of individual species in
monoculture. We then confront our theoretical predictions with
simulations from the empirically-based population models. Second, we use
the multi-species population models to perform simulation experiments
that isolate the effects of environmental stochasticity, demographic
stochasticity, and interspecific \textcolor{blue}{interactions} on
community-wide species synchrony. Given that our population models
capture the essential features of community dynamics important to
synchrony (density-dependence, and demographic and environmental
stochasticity), and that these models successfully reproduce observed
community dynamics (Chu and Adler 2015), perturbing the models can
reveal the processes that determine species synchrony in our focal
grassland communities.

\subsection{THEORETICAL MODEL}\label{theoretical-model}

\subsubsection{The model}\label{the-model}

While existing theory has identified the factors driving synchrony, we
do not have a simple expression to predict synchrony in a particular
community with all factors operating simultaneously. However, we can
derive analytical predictions for species synchrony under special
limiting cases.
\textcolor{blue}{The limiting case predictions we derive serve as baselines to help us interpret results from empirically-based simulations (described below).}
We focus on synchrony of per capita growth rates, rather than
abundances, because growth rates represent the instantaneous response of
species to the environment and competition, and are less susceptible to
the legacy effects of drift and disturbance (Loreau and {{de
Mazancourt}} 2008). We present equivalent results for synchrony of
species abundances in the Online Supporting Information, and show that
they lead to the same overall conclusions as synchrony of per capita
growth rates. Following Loreau and {{de Mazancourt}} (2008) and {{de
Mazancourt}} et al. (2013), we define population growth, ignoring
observation error, as

\begin{align}
r_i(t) &= \text{ln}N_i(t+1) - \text{ln}N_i(t) \\
&= r_{mi} \left[ 1- \frac{N_i(t)+\sum_{j \neq i} \alpha_{ij}N_j(t)} {K_i} + \sigma_{ei}u_{ei}(t) + \frac{\sigma_{di}u_{di}(t)}{\sqrt{N_i(t)}} \right]
\end{align}

\noindent where \(N_i(t)\) is the biomass of species \emph{i} in year
\emph{t}, and \(r_i(t)\) is its population growth rate in year \emph{t}.
\(r_{mi}\) is species \emph{i}'s intrinsic rate of increase, \(K_i\) is
its carrying capacity, and \(\alpha_{ij}\) is the interspecific
competition coefficient representing the effect of species \emph{j} on
species \emph{i}. Environmental stochasticity is incorporated as
\(\sigma_{ei}u_{ei}(t)\), where \(\sigma_{ei}^2\) is the environmental
variance and \(u_{ei}\) are normal random variables with zero mean and
unit variance that are independent through time but may be correlated
between species. Demographic stochasticity arises from variations in
births and deaths among individuals (e.g., same states, different
fates), and is included in the model as a first-order, normal
approximation (Lande et al. 2003, {{de Mazancourt}} et al. 2013).
\(\sigma_{di}^2\) is the demographic variance and \(u_{di}(t)\) are
independent normal variables with zero mean and unit variance. To derive
analytical predictions we solved a first-order approximation of Equation
2 ({{de Mazancourt}} et al. 2013 and Online Supporting Information).
\textcolor{blue}{Due to the linear approximation approach, our analytical predictions will likely fail in communities where species exhibit large fluctuations due to limit cycles and chaos}
(Loreau and {{de Mazancourt}} 2008).
\textcolor{blue}{Indeed, one of the advantages of focusing on growth rates rather than abundances is that growth rates are more likely to be well-regulated around an equilibrium value, if the long-term average of a species' growth rate is relatively small}
(e.g., \(r < 2\)).

\subsubsection{Predictions}\label{predictions}

Our first prediction assumes no interspecific interactions, no
environmental stochasticity, identical intrinsic growth rates, and that
demographic stochasticity is operating but all species have identical
demographic variances. This limiting case, \(\mathcal{M}_{D}\),
represents a community where dynamics are driven by demographic
stochasticity alone. Our prediction
\textcolor{blue}{for the synchrony of per capita growth rates for $\mathcal{M}_{D}$, $\phi_{R,\mathcal{M}_{D}}$, is}

\begin{equation}
\phi_{R,\mathcal{M}_{D}} = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2},
\end{equation}

\noindent where \(p_i\) is the average frequency of species \emph{i},
\(p_i = N_i/N_T\). When all species have identical abundances and
\(p_i = 1/S\), where \emph{S} is species richness, synchrony equal 1/S
(Loreau and {{de Mazancourt}} 2008).

Our second limiting case assumes only environmental stochasticity is
operating (\(\mathcal{M}_{E}\)). Thus, we assume there are no
interspecific interactions, demographic stochasticity is absent,
intrinsic growth rates are identical, and environmental variance is
identical for all species. Our prediction
\textcolor{blue}{for the synchrony of per capita growth rates for $\mathcal{M}_{E}$, $\phi_{R,\mathcal{M}_{E}}$, is }

\begin{equation}
\phi_{R,\mathcal{M}_{E}} = \frac{\sum_{i,j}\text{cov}(u_{ei},u_{ej})}{S^2},
\end{equation}

\noindent{} where \(\text{cov}(u_{ei},u_{ej})\) is the standardized
covariance of environmental responses between species \emph{i} and
species \emph{j}. Confronting our theoretical predictions with data
requires estimates of species dynamics of large populations (no
demographic stochasticity) \textcolor{blue}{growing in isolation} (no
interspecific interactions) to calculate the covariance of species'
environmental responses. To estimate environmental responses in natural
communities, we turn to our population models built using long-term
demographic data.

\subsection{EMPIRICAL ANALYSIS}\label{empirical-analysis}

\subsubsection{Materials and methods}\label{materials-and-methods}

\paragraph{Data}\label{data}

We use long-term demographic data from five semiarid grasslands in the
western United States (described in detail by Chu and Adler 2015). Each
site includes a set of 1-\(\text{m}^2\) permanent quadrats within which
all individual plants were identified and mapped annually using a
pantograph (Hill 1920). The resulting mapped polygons represent basal
cover for grasses and canopy cover for shrubs. Data come from the
Sonoran desert in Arizona (Anderson et al. 2012), sagebrush steppe in
Idaho (Zachmann et al. 2010), southern mixed prairie in Kansas (Adler et
al. 2007), northern mixed prairie in Montana (Anderson et al. 2011), and
Chihuahuan desert in New Mexico (Anderson et al. in preparation, Chu and
Adler 2015) (Table 1).

\paragraph{Calculating observed
synchrony}\label{calculating-observed-synchrony}

The data consist of records for individual plant size in quadrats for
each year. To obtain estimates of percent cover for each focal species
in each year, we summed the individual-level data within quadrats and
then averaged percent cover, by species, over all quadrats. We
calculated per capita growth rates as
\(\text{log}(x_t) - \text{log}(x_{t-1})\), where \emph{x} is species'
percent cover in year \emph{t}. Using the community time series of per
capita growth rates or percent cover, we calculated community synchrony
using the metric of Loreau and {{de Mazancourt}} (2008) in the
`synchrony' package (Gouhier and Guichard 2014) in R (R Core Team 2013).
Specifically, we calculated synchrony as

\begin{equation}
\phi_r = \frac{\sigma^{2}_{T}}{(\sum_{i}\sigma_{r_{i}})^{2}}
\end{equation}

\noindent where \(\sigma_{r_{i}}\) is the temporal variance of species
\emph{i'}s per capita population growth rate (\(r_i\)) and
\(\sigma^{2}_{T}\) is the temporal variance of the aggregate
community-level growth rate. \(\phi\) ranges from 0 at perfect
asynchrony to 1 at perfect synchrony (Loreau and {{de Mazancourt}}
2008). We use the same equation to calculate observed synchrony of
species' percent cover, which we present to relate our results to
previous findings, even though we focus on synchrony of growth rates in
our model simulations (see below).

\paragraph{Fitting statistical models}\label{fitting-statistical-models}

Vital rate regressions are the building blocks of our dynamic models: an
integral projection model (IPM) and an individual-based model (IBM). We
followed the approach of Chu and Adler (2015) to fit statistical models
for survival, growth, and recruitment (see Online Supporting Information
for full details). We modeled survival probability of each genet as
function of genet size, temporal variation among years, permanent
spatial variation among groups of quadrats, and local neighborhood
crowding from conspecific and heterospecific genets. Regression
coefficients for the effect of crowding by each species can be
considered a matrix of interaction coefficients whose diagonals
represent intraspecific interactions and whose off-diagonals represent
interspecific interactions (Adler et al. 2010). These interaction
coefficients can take positive (facilitative) or negative (competitive)
values. We modeled growth as the change in size of a genet from one year
to the next, which depends on the same factors as the survival model. We
fit the survival and growth regressions using INLA (Rue et al. 2014), a
statistical package for fitting generalized linear mixed effects models
via approximate Bayesian inference (Rue et al. 2009), in R (R Core Team
2013).
\textcolor{blue}{Crowding was treated as a fixed effect without a temporal component because most 95\% credible intervals for random year effects on crowding broadly overlapped zero and, in a test case, including yearly crowding effects did not change our results.}
Spatial (quadrat groupings) variation was treated as a random effect on
the intercept and temporal (interannual) variation was treated as random
effects on the intercept and the effect of genet size in the previous
year (Online Supporting Information).

We modeled recruitment at the quadrat scale, rather than the individual
scale, because the original data do not attribute new genets to specific
parents (Chu and Adler 2015). Our recruitment model assumes that the
number of recruits produced in each year follows a negative binomial
distribution with the mean dependent on the cover of the parent species,
permanent spatial variation among groups, temporal variation among
years, and inter- and intraspecific interactions as a function of total
species' cover in the quadrat. We fit the recruitment model using a
hierarchical Bayesian approach implemented in JAGS (Plummer 2003) using
the `rjags' package (Plummer 2014) in R (R Core Team 2013). Again,
temporal and spatial variation were treated as random effects.

\paragraph{Building dynamic multi-species
models}\label{building-dynamic-multi-species-models}

Once we have fit the vital rate statistical models, building the
population models is straightforward. For the IBM, we initialize the
model by randomly assigning plants spatial coordinates, sizes, and
species identities until each species achieves a density representative
of that observed in the data. We then project the model forward by using
the survival regression to determine whether a genet lives or dies, the
growth regression to calculate changes in genet size, and the
recruitment regression to add new individuals that are distributed
randomly in space. Crowding is directly calculated at each time step
since each genet is spatially referenced (as in the observed data).
Environmental stochasticity is not an inherent feature of IBMs, but is
easily included since we fit year-specific temporal random effects for
each vital rate regression. To include temporal environmental variation,
at each time step we randomly choose a set of estimated survival,
growth, and recruitment parameters specific to one observation year. For
all simulations, we ignore the spatial random effect associated with
variation among quadrat groups, so our simulations represent an average
quadrat for each site.

The IPM uses the same vital rate regressions as the IBM (Rees and Ellner
2009, Rees et al. 2014), but it is spatially implicit and does not
include demographic stochasticity. Following Chu and Adler (2015), we
use a mean field approximation that captures the essential features of
spatial patterning to define the crowding index at each time step
(Supporting Online Information). Temporal variation is included in
exactly the same way as for the IBM. For full details on the IPMs we
use, see Chu and Adler (2015).

\paragraph{Simulation experiments}\label{simulation-experiments}

We performed simulation experiments where drivers (environmental
stochasticity, demographic stochasticity, or interspecific
\textcolor{blue}{interactions}) were removed one-by-one and in
combination. To remove interspecific interactions, we set the
off-diagonals of the interaction matrix for each vital rate regression
to zero. This retains intraspecific interactions, and thus
density-dependence, and results in simulations where species are
\textcolor{blue}{growing in isolation}.
\textcolor{blue}{We cannot definitively rule out the effects of species interactions on all parameters, meaning that a true monoculture could behave differently than our simulations of a population growing without interspecific competitors.}
To remove the effect of a fluctuating environment, we removed the
temporal (interannual) random effects from the regression equations. To
remove the effect of demographic stochasticity, we use the IPM rather
than the IBM because the IPM does not include demographic stochasticity
(demographic stochasticity cannot be removed from the IBM). Since the
effect of demographic stochasticity on population dynamics depends on
population size (Lande et al. 2003), we can control the strength of
demographic stochasticity by simulating the IBM on areas (e.g.~plots) of
different size. Indeed, results from an IBM with infinite population
size would converge on results from the IPM. Given computational
constraints, the largest landscape we simulate is a 25 \(\text{m}^2\)
plot.

We conducted the following six simulation experiments: (1) IBM: All
drivers (environmental stochasticity, demographic stochasticity, or
\textcolor{blue}{interspecific interactions}) present; (2) IPM:
Demographic stochasticity removed; (3) IBM: Environmental stochasticity
removed; (4) IBM: \textcolor{blue}{Interspecific interactions} removed;
(5) IPM: \textcolor{blue}{Interspecific interactions} and demographic
stochasticity removed; (6) IBM:
\textcolor{blue}{Interspecific interactions} and environmental
stochasticity removed. We ran IPM simulations for 2,000 time steps,
after an initial 500 iteration burn-in period. This allowed species time
to reach their stable size distribution. We then calculated the
synchrony of species' per capita growth rates over 100 randomly selected
contiguous 50 time-step sections. We ran IBM simulations for 100 time
steps, and repeated the simulations 100 times for each simulation
experiment. From those, we retained only the simulations in which no
species went extinct due to demographic stochasticity. Synchrony was
calculated over the 100 time steps for each no extinction run within a
model experiment. To explore the effect of demographic stochasticity in
different sized populations, we ran simulations (1) and (6) on plot
sizes of 1, 4, 9, 16, and 25 \(\text{m}^2\). All other IBM simulations
were run on a 25 \(\text{m}^2\) landscape.

Results from our simulation experiments also allow us to test our
theoretical predictions. First, in the absence of
\textcolor{blue}{interspecific interactions} and demographic
stochasticity,
\textcolor{blue}{populations can only fluctuate in response to the environment}.
Therefore, we can use results from simulation (5) to estimate the
covariance of species' responses to the environment
(\(\text{cov}(u_{ie}, u_{je})\)) and parameterize Equation 4.
Parameterizing Equation 3 does not require simulation output because the
only parameters are the species' relative abundances. Second,
simulations (5) and (6) represent the simulated version of our limiting
case theoretical predictions. Thus, we directly test the theoretical
predictions by comparing them to observed synchrony and simulated
synchrony.

\subsubsection{Results}\label{results}

Synchrony of species' per capita growth rates at our study sites range
from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to
0.92 (Table 2). Synchrony tends to be higher in communities with few
species (Arizona and New Mexico) and/or with relatively high temporal
variability in percent cover (e.g., Montana). Synchrony is lowest in
Idaho, the only data set that includes two life forms: a shrub and three
perennial grasses (Table 1).
\textcolor{blue}{Synchrony of per capita growth rates and CV of percent cover are positively correlated}
(Pearson's \(\rho\) = 0.72). For all five communities, species synchrony
from IPM and IBM simulations closely approximated observed synchrony
(Fig. S1). IBM-simulated synchrony is consistently, but only slightly,
lower than IPM-simulated synchrony (Fig. S1), likely due to the
desynchronizing effect of demographic stochasticity.

Across the five communities, our limiting case predictions closely
matched synchrony from the corresponding simulation experiment (Fig. 1
and Table S1). The correlation between our analytical predictions and
simulated synchrony was 0.97 for \(\phi_{R,\mathcal{M}_D}\) and 0.997
for \(\phi_{R,\mathcal{M}_E}\). The largest difference between predicted
and simulated synchrony was 0.05 in New Mexico for
\(\phi_{R,\mathcal{M}_D}\) (Table S1).

Simulation experiments revealed that removing environmental fluctuations
has the largest impact on synchrony, leading to a reduction in synchrony
of species growth rates in four out of five communities (Fig. 1).
Removing environmental fluctuations (``No E.S'' simulations) decreased
synchrony by 33\% in Arizona, 48\% in Kansas, 39\% in Montana, and 40\%
in New Mexico. Only in Idaho did removing environmental fluctuations
cause an increase in synchrony (Fig. 1), but the effect was small (9\%
increase; Table S2). Overall, species' temporal random effects in the
statistical vital rate models are positively correlated (Table S3).
Species interactions are weak in these communities (Table S4 and Chu and
Adler 2015), and removing \textcolor{blue}{interspecific interactions}
had little effect on synchrony (Fig. 1; ``No Comp.'' simulations).
Removing \textcolor{blue}{interspecific interactions} caused, at most, a
5\% change in synchrony (Fig. 1). Removing demographic stochasticity
(``No D.S.'' simulations) caused synchrony to increase slightly in all
communities (Fig. 1), with an average 6\% increase over synchrony from
IBM simulations on a 25\(\text{m}^2\) area.

The desynchronizing effect of demographic stochasticity, which increases
as population size decreases, modestly counteracted the synchronizing
force of the environment, but not enough to lower synchrony to the level
observed when only demographic stochasticity is operating (Fig. 2). In
the largest, 25 \(\text{m}^2\) plots, synchrony was driven by
environmental stochasticity (e.g., \(\mathcal{M}_E\)). At 1
\(\text{m}^2\), synchrony reflected demographic stochasticity and
environmental stochasticity (e.g., between \(\mathcal{M}_E\) and
\(\mathcal{M}_D\)). For context, population sizes increased from an
average of 17 individuals per community in 1 \(\text{m}^2\) IBM
simulations to an average of 357 individuals per community in 25
\(\text{m}^2\) IBM simulations.

For all five communities, the synchrony of species' growth rates when
growing in isolation almost perfectly matched species synchrony in
polyculture (Fig. 3). Results for synchrony of percent cover are
qualitatively similar, but simulation results were more variable and
less consistent with analytical predictions and observed synchrony
(Online Supporting Information, Figs. S2-S3).

\subsection{DISCUSSION}\label{discussion}

Our study produced four main findings that were generally consistent
across five natural plant communities: (1) limiting-case predictions
from the theoretical model were well-supported by simulations from the
empirical models; (2) demographic stochasticity decreased community
synchrony, as expected by theory, and its effect was largest in small
populations; (3) environmental fluctuations increased community
synchrony relative to simulations in constant environments because
species-specific responses to the environment were positively, though
not perfectly, correlated; and (4) interspecific interactions were weak
and therefore had little impact on community synchrony. We also found
that analyses based on synchrony of species' percent cover, rather than
growth rates, were uninformative (Figs. S2-S3) since the linear
approximation required for analytical predictions is a stronger
assumption for abundance than growth rates, especially given relatively
short time-series (Online Supporting Information). Thus, our results
provide further evidence that it is difficult to decipher mechanisms of
species synchrony from abundance time series, as expected by theory
(Loreau and {{de Mazancourt}} 2008).
\textcolor{blue}{Observed synchrony of per capita growth rates was positively correlated with the variability of percent cover across our focal communities, which confirms that we are investigating an important process underlying ecosystem stability.}

\subsubsection{Simulations support theoretical
predictions}\label{simulations-support-theoretical-predictions}

Our theoretical predictions were derived from a simple model of
population dynamics and required several simplifying assumptions,
raising questions about their relevance to natural communities.
\textcolor{blue}{For example, the species in our communities do not have equivalent environmental and demographic variances (Figs. S4-S7), as required by our predictions.}
However, the theoretical predictions
\textcolor{blue}{closely matched results from} simulations of population
models fit to long-term data from natural plant communities (Table 3).
Such \textcolor{blue}{strong agreement} between our analytical
predictions and the simulation results should inspire confidence in the
ability of simple models to inform our understanding of species
synchrony
\textcolor{blue}{even in complex natural communities, and allows us to place our simulation results within the context of contemporary theory}.

\subsubsection{Demographic stochasticity decreases
synchrony}\label{demographic-stochasticity-decreases-synchrony}

Demographic stochasticity partially counteracted the synchronizing
effects of environmental fluctuations and interspecific interactions on
per capita growth rates, but only when populations were small (Fig. 2),
in agreement with theory (Loreau and {{de Mazancourt}} 2008). Even in
small populations, however, demographic stochasticity was not strong
enough to \textcolor{blue}{compensate} the synchronizing effects of
environmental fluctuations and match the analytical prediction where
only demographic stochasticity is operating (\(\mathcal{M}_D\) in Fig.
2). These results confirm the theoretical argument by Loreau and {{de
Mazancourt}} (2008) that independent fluctuations among interacting
species in a non-constant environment should be rare. Only in the Idaho
community does synchrony of per capita growth rates approach
\(\mathcal{M}_D\) in a non-constant environment (Fig. 2). This is most
likely due to the strong effect of demographic stochasticity on the
shrub \emph{Artemisia tripartita} since even a \textcolor{blue}{2}5
\(\text{m}^2\) quadrat would only contain a few individuals of that
species.

Our analysis of how demographic stochasticity affects synchrony
demonstrates that synchrony depends on the observation area. As the
observation area increases, population size increases and the
desynchronizing effect of demographic stochasticity lessens (Fig. 2).
Thus, our results suggest that community-wide species synchrony will
increase as the observation area increases, rising from
\(\mathcal{M}_D\) to \(\mathcal{M}_E\). Such a conclusion assumes,
however, that species richness remains constant as observation area
increases, which is unlikely (Taylor 1961). Recent theoretical work has
begun to explore the linkage between ecosystem stability and spatial
scale (Wang and Loreau 2014, 2016), and our results suggest that
including demographic stochasticity in theoretical models of
metacommunity dynamics may be important for understanding the role of
species synchrony in determining ecosystem stability across spatial
scales.

\subsubsection{Environmental fluctuations drive community
synchrony}\label{environmental-fluctuations-drive-community-synchrony}

In large populations where interspecific interactions are weak,
synchrony is expected to be driven exclusively by environmental
fluctuations (Equation 4). Under such conditions community synchrony
should approximately equal the synchrony of species' responses to the
environment (Loreau and {{de Mazancourt}} 2008). Two lines of evidence
lead us to conclude that environmental fluctuations drive species
synchrony in our focal plant communities. First, in our simulation
experiments, removing interspecific interactions resulted in no
discernible change in community-wide species synchrony of per capita
growth rates (Fig. 1). Second, removing environmental fluctuations from
simulations consistently reduced synchrony (Fig. 1), and the synchrony
of species in isolation was a very strong predictor of synchrony of
species in polyculture (Fig. 3). Our results lead us to conclude that
environmental fluctuations, not species interactions, are the primary
driver of community-wide species synchrony in the communities we
studied. Given accumulating evidence that niche differences in natural
communities are large (reviewed in Chu and Adler 2015), and thus species
interactions are likely to be weak, our results may be general in
natural plant communities.

In the Idaho community, removing environmental fluctuations did not
cause a large decrease in synchrony. However, that result appears to be
an artifact. Removing environmental variation results in a negative
invasion growth rate for \emph{A. tripartita}. Although we only analyzed
IBM runs in which \emph{A. tripartita} had not yet gone extinct, it was
at much lower abundance than in the other simulation runs. When we
removed \emph{A. tripartita} from all simulations, the Idaho results
conformed with results from all other sites: removing environmental
stochasticity caused a significant reduction in species synchrony (Fig.
S8). Our main results for Idaho (Fig. 2), with \emph{A. tripartita}
included, demonstrate how the processes that determine species synchrony
interact in complex ways. \emph{A. tripartita} has a facilitative effect
on each grass species across all vital rates, except for a small
competitive effect on \emph{H. comata}'s survival probability (Tables
S8-S10). At the same time, all the perennial grasses have negative
effects on each other for each vital rate (Tables S8-S10). We know
synchrony is affected by interspecific competition (Loreau and {{de
Mazancourt}} 2008), but how facilitative effects manifest themselves is
unknown. The interaction of facilitation and competition is clearly
capable of having a large effect on species synchrony, and future
theoretical efforts should aim to include a wider range of species
interactions.

A challenge to the generality of our results is that we were only able
to model common, co-occuring species (see Chu and Adler 2015).
\textcolor{blue}{Most communities are dominated by few common species and many rare species}
(McGill et al. 2007),
\textcolor{blue}{meaning that the low number of common species in our focal commiunities is not unusual.}
\textcolor{blue}{Rather, the generality of our results hinges upon the influence of rare species.
Rare species could be kept at low densities by competitive suppression due to strong interspecific interactions.
If we had been able to model rare species, we might have found a larger effect of competition on community synchrony.
However, recent theory predicts that persistent rare species may actually experience even weaker interspecific competition, and thus stronger niche differences, than common species}
(Yenni et al. 2012),
\textcolor{blue}{in which case synchrony would remain predominantly driven by environmental responses.}
Rare species could also be limited by generalist natural enemies whose
effects are density-independent but alter competitive hierarchies
(Mordecai 2011). Under such conditions, synchrony will depend on the
degree of pathogen-induced fitness differences and the pathogen's
response to environmental conditions. Neither our model nor current
theory explicitly considers the effects of pathogens on species
synchrony, and this highlights the need for theoretical work on the
interaction between mechanisms of species coexistence and mechanisms of
species synchrony (Loreau 2010).

\subsubsection{Interspecific interactions had little impact on community
synchrony}\label{interspecific-interactions-had-little-impact-on-community-synchrony}

We expected community synchrony of per capita growth rates to decrease
when we removed interspecific interactions (Loreau and {{de Mazancourt}}
2008). We found that community synchrony was virtually indistinguishable
between simulations with and without interspecific interactions (Fig.
2). The lack of an effect of
\textcolor{blue}{interspecific interactions} on synchrony is in contrast
to a large body of theoretical work that predicts a strong role for
competition in creating compensatory dynamics (Tilman 1988) and a recent
empirical analysis (Gross et al. 2014).

Our results do not contradict the idea that competition can lead to
compensatory dynamics, but they do highlight the fact that interspecific
competition must be relatively strong to influence species synchrony.
The communities we analyzed are composed of species with very little
niche overlap (Chu and Adler 2015) and weak interspecific interactions
(Tables S1, S3-S17). Mechanistic consumer-resource models (Lehman and
Tilman 2000) and phenomenological Lotka-Volterra models (Lehman and
Tilman 2000, Loreau and {{de Mazancourt}} 2013) both confirm that the
effect of competition on species synchrony diminishes as niche overlap
decreases. In that sense, our results are not surprising: interspecific
interactions are weak, so of course removing them does not affect
synchrony.

However, our results do contrast with a recent analysis of several
biodiversity-ecosystem functioning experiments showing that competition
drives species synchrony in grasslands (Gross et al. 2014). The apparent
inconsistency between our results and those of Gross et al. (2014) may
be explained by the differences between our studies. Gross et al.'s
results are based on rapidly assembling experimental communities that
are, at most, 11 years old. The natural, relatively undisturbed
communities we studied consist of species that have co-occurred for
decades (Chu and Adler 2015) and represent a much later stage of
community assembly. In theory, species interactions should weaken
through time as community assembly proceeds (Kokkoris et al. 1999),
meaning that in biodiversity-ecosystem functioning experiments the
processes driving competitive exclusion are still operating. Such
communities may be highly influenced by transient, but strong,
interspecific competition that can mask the synchronizing effects of the
environment. In contrast, synchrony may be driven by environmental
fluctuations in older communities. In other words, the dominant driver
of community synchrony should shift from competition to environmental
fluctuations through time. One way to test this prediction is to
continue collecting data from biodiversity-ecoystem functioning
experiments and conduct the analysis of Gross et al. (2014) every few
years.

\textcolor{blue}{Our conclusion that species interactions have little impact on synchrony only applies to single trophic level communities.
Species interactions almost certainly play a strong role in multi-trophic communities where factors such as resource overlap}
(Vasseur and Fox 2007), \textcolor{blue}{dispersal} (Gouhier et al.
2010), \textcolor{blue}{and the strength of top-down control} (Bauer et
al. 2014)
\textcolor{blue}{are all likely to affect community synchrony.}

\subsection{CONCLUSIONS}\label{conclusions}

Species-specific responses to temporally fluctuating environmental
conditions is an important mechanism underlying asynchronous population
dynamics and, in turn, ecosystem stability (Loreau and {{de Mazancourt}}
2013). When we removed environmental variation, we found that synchrony
decreased in four out of the five grassland communities we studied (Fig.
2). A tempting conclusion is that our study confirms that compensatory
dynamics are rare in natural communities, and that ecologically-similar
species will exhibit synchronous dynamics (e.g., Houlahan et al. 2007).
Such a conclusion misses an important subtlety. The perennial grasses we
studied do have similar responses to the environment (Table S2), which
will tend to synchronize dynamics. However, if community-wide species
synchrony is less than 1, as it is in all our focal communities, some
degree of compensatory dynamics must be present (Loreau and {{de
Mazancourt}} 2008). In agreement with other studies (Rocha et al. 2011,
Vasseur et al. 2014), we find that environmental responses are primarily
responsible for the degree of synchrony among ecologically-similar
species. This result contrasts with a recent analysis of several
biodiversity-ecosystem functioning experiments showing that competition
drives community synchrony (Gross et al. 2014). Recently assembled
communities, such as experimental plots, may exhibit strong species
interactions that will weaken over time (Kokkoris et al. 1999). Future
research on the influence of community assembly on synchrony could
reconcile inconsistent observations on the biotic and abiotic forces
that jointly drive community dynamics.

\subsection{ACKNOWLEDGMENTS}\label{acknowledgments}

This work was funded by the National Science Foundation through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
CAREER award to PBA (DEB-1054040). CdM and ML were supported by the
TULIP Laboratory of Excellence (ANR-10-LABX-41). We thank the original
mappers of the permanent quadrats at each site and the digitizers in the
Adler lab, without whom this work would not have been possible. Compute,
storage, and other resources from the Division of Research Computing in
the Office of Research and Graduate Studies at Utah State University are
gratefully acknowledged. We also thank three anonymous reviewers who
provided critical feedback that improved the manuscript.

\setlength{\parindent}{0ex}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
perennial plants: An embarrassment of niches. Ecology Letters
13:1019--1029.

Adler, P. B., W. R. Tyburczy, and W. K. Lauenroth. 2007. Long-term
mapped quadrats from Kansas prairie: demographic information for
herbaceous plants. Ecology 88:2673.

Anderson, J., M. P. McClaran, and P. B. Adler. 2012. Cover and density
of semi-desert grassland plants in permanent quadrats mapped from 1915
to 1947. Ecology 93:1492--1492.

Anderson, J., L. Vermeire, and P. B. Adler. 2011. Fourteen years of
mapped, permanent quadrats in a northern mixed prairie, USA. Ecology
92:1703.

Bauer, B., M. Vos, T. Klauschies, and U. Gaedke. 2014. Diversity,
Functional Similarity, and Top-Down Control Drive Synchronization and
the Reliability of Ecosystem Function. The American Naturalist
183:394--409.

Cardinale, B. J., J. E. Duffy, A. Gonzalez, D. U. Hooper, C. Perrings,
P. Venail, A. Narwani, G. M. Mace, D. Tilman, D. A. Wardle, A. P.
Kinzig, G. C. Daily, M. Loreau, and J. B. Grace. 2012. Biodiversity loss
and its impact on humanity. Nature 489:59--67.

Chu, C., and P. B. Adler. 2015. Large niche differences emerge at the
recruitment stage to stabilize grassland coexistence. Ecological
Monographs 85:373--392.

{{de Mazancourt}}, C., F. Isbell, A. Larocque, F. Berendse, E. {De
Luca}, J. B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B.
Schmid, D. Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M.
Loreau. 2013. Predicting ecosystem stability from community composition
and biodiversity. Ecology Letters 16:617--625.

Dornelas, M., N. J. Gotelli, B. McGill, H. Shimadzu, F. Moyes, C.
Sievers, and A. E. Magurran. 2014. Assemblage time series reveal
biodiversity change but not systematic loss. Science 344:296--9.

Gonzalez, A., and M. Loreau. 2009. The Causes and Consequences of
Compensatory Dynamics in Ecological Communities. Annual Review of
Ecology, Evolution, and Systematics 40:393--414.

Gouhier, T. C., and F. Guichard. 2014. Synchrony: quantifying
variability in space and time. Methods in Ecology and Evolution
5:524--533.

Gouhier, T. C., F. Guichard, and A. Gonzalez. 2010. Synchrony and
Stability of Food Webs in Metacommunities. The American Naturalist
175:E16--E34.

Gross, K., B. J. Cardinale, J. W. Fox, A. Gonzalez, M. Loreau, H. W.
Polley, P. B. Reich, and J. van Ruijven. 2014. Species richness and the
temporal stability of biomass production: a new analysis of recent
biodiversity experiments. The American naturalist 183:1--12.

Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
of diversity in natural grasslands. Nature 508:521--5.

Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
diversity on grassland productivity through population asynchrony and
overyielding. Ecology 91:2213--2220.

Hooper, D., F. {Chapin III}, and J. Ewel. 2005. Effects of biodiversity
on ecosystem functioning: a consensus of current knowledge. Ecological
Monographs 75:3--35.

Houlahan, J. E., D. J. Currie, K. Cottenie, G. S. Cumming, S. K. M.
Ernest, C. S. Findlay, S. D. Fuhlendorf, U. Gaedke, P. Legendre, J. J.
Magnuson, B. H. McArdle, E. H. Muldavin, D. Noble, R. Russell, R. D.
Stevens, T. J. Willis, I. P. Woiwod, and S. M. Wondzell. 2007.
Compensatory dynamics are rare in natural ecological communities.
Proceedings of the National Academy of Sciences of the United States of
America 104:3273--3277.

Isbell, F. I., H. W. Polley, and B. J. Wilsey. 2009. Biodiversity,
productivity and the temporal stability of productivity: Patterns and
processes. Ecology Letters 12:443--451.

Kokkoris, G. D., A. Y. Troumbis, and J. H. Lawton. 1999. Patterns of
species interaction strength in assembled theoretical competition
communities. Ecology Letters 2:70--74.

Lande, R., S. Engen, and B.-E. Saether. 2003. Stochastic population
dynamics in ecology and conservation.

Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
Productivity in Competitive Communities. The American Naturalist
156:534--552.

Loreau, M. 2010. From Populations to Ecosystems: Theoretical Fondations
for a New Ecological Synthesis.

Loreau, M., and C. {{de Mazancourt}}. 2008. Species synchrony and its
drivers: neutral and nonneutral community dynamics in fluctuating
environments. The American Naturalist 172:E48--E66.

Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
stability: A synthesis of underlying mechanisms. Ecology Letters
16:106--115.

Loreau, M., S. Naeem, P. Inchausti, J. Bengtsson, J. P. Grime, A.
Hector, D. U. Hooper, M. A. Huston, D. Raffaelli, B. Schmid, D. Tilman,
and D. A. Wardle. 2001. Biodiversity and ecosystem functioning: current
knowledge and future challenges. Science 294:804--808.

McGill, B. J., R. S. Etienne, J. S. Gray, D. Alonso, M. J. Anderson, H.
K. Benecha, M. Dornelas, B. J. Enquist, J. L. Green, F. He, A. H.
Hurlbert, A. E. Magurran, P. A. Marquet, B. A. Maurer, A. Ostling, C. U.
Soykan, K. I. Ugland, and E. P. White. 2007. Species abundance
distributions: Moving beyond single prediction theories to integration
within an ecological framework. Ecology Letters 10:995--1015.

Mordecai, E. A. 2011. Pathogen impacts on plant communities: Unifying
theory, concepts, and empirical work. Ecological Monographs 81:429--441.

Mutshinda, C. M., R. B. O'Hara, and I. P. Woiwod. 2009. What drives
community dynamics? Proceedings of the Royal Society B: Biological
Sciences 276:2923--2929.

Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
3rd international workshop on distributed statistical computing (dSC
2003). march.

Plummer, M. 2014. rjags: Bayesian graphical models using MCMC.

R Core Team. 2013. R: A language and environment for statistical
computing.

Ranta, E., V. Kaitala, M. S. Fowler, J. Laakso, L. Ruokolainen, and R.
O'Hara. 2008. Detecting compensatory dynamics in competitive communities
under environmental forcing. Oikos 117:1907--1911.

Rees, M., and S. P. Ellner. 2009. Integral projection models for
populations in temporally varying environments. Ecological Monographs
79:575--594.

Rees, M., D. Z. Childs, and S. P. Ellner. 2014. Building integral
projection models: A user's guide.

Rocha, M. R., U. Gaedke, and D. a. Vasseur. 2011. Functionally similar
species have similar dynamics. Journal of Ecology 99:1453--1459.

Rockstrom, J., J. Rockstrom, W. Steffen, W. Steffen, K. Noone, K. Noone,
A. Persson, A. Persson, F. S. Chapin, F. S. Chapin, E. F. Lambin, E. F.
Lambin, T. M. Lenton, T. M. Lenton, M. Scheffer, M. Scheffer, C. Folke,
C. Folke, H. J. Schellnhuber, H. J. Schellnhuber, B. Nykvist, B.
Nykvist, C. A. de Wit, C. A. de Wit, T. Hughes, T. Hughes, S. van der
Leeuw, S. van der Leeuw, H. Rodhe, H. Rodhe, S. Sorlin, S. Sorlin, P. K.
Snyder, P. K. Snyder, R. Costanza, R. Costanza, U. Svedin, U. Svedin, M.
Falkenmark, M. Falkenmark, L. Karlberg, L. Karlberg, R. W. Corell, R. W.
Corell, V. J. Fabry, V. J. Fabry, J. Hansen, J. Hansen, B. Walker, B.
Walker, D. Liverman, D. Liverman, K. Richardson, K. Richardson, P.
Crutzen, P. Crutzen, J. A. Foley, and J. A. Foley. 2009. A safe
operating space for humanity. Nature 461:472--475.

Rue, H., S. Martino, and N. Chopin. 2009. Approximate Bayesian Inference
for Latent Gaussian Models Using Integrated Nested Laplace
Approximations. Journal of the Royal Statistical Society, Series B
71:319--392.

Rue, H., S. Martino, F. Lindgren, D. Simpson, A. Riebler, and E.
Teixeira. 2014. INLA: Functions which allow to perform full Bayesian
analysis of latent Gaussian models using Integrated Nested Laplace
Approximaxion.

Taylor, L. R. 1961. Aggregation, Variance and the Mean. Nature
189:732--735.

Thibaut, L. M., S. R. Connolly, and H. P. A. Sweatman. 2012. Diversity
and stability of herbivorous fishes on coral reefs. Ecology 93:891--901.

Tilman, D. 1988. Plant strategies and the dynamics and structure of
plant communities. Plant strategies and the dynamics and structure of
plant communities.:360 pp.

Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
ecosystem stability in a decade-long grassland experiment. Nature
441:629--632.

Tilman, D., P. B. Reich, J. Knops, D. Wedin, T. Mielke, and C. Lehman.
2001. Diversity and productivity in a long-term grassland experiment.
Science 294:843--845.

Vasseur, D. a, J. W. Fox, A. Gonzalez, R. Adrian, B. E. Beisner, M. R.
Helmus, C. Johnson, P. Kratina, C. Kremer, C. de Mazancourt, E. Miller,
W. a Nelson, M. Paterson, J. a Rusak, J. B. Shurin, and C. F. Steiner.
2014. Synchronous dynamics of zooplankton competitors prevail in
temperate lake ecosystems. Proceedings. Biological sciences / The Royal
Society 281:20140633.

Vasseur, D. A., and J. W. Fox. 2007. Environmental fluctuations can
stabilize food web dynamics by increasing synchrony. Ecology Letters
10:1066--1074.

Vellend, M., L. Baeten, I. H. Myers-Smith, S. C. Elmendorf, R.
Beaus{é}jour, C. D. Brown, P. {De Frenne}, K. Verheyen, and S. Wipf.
2013. Global meta-analysis reveals no net change in local-scale plant
biodiversity over time. Proceedings of the National Academy of Sciences
of the United States of America 110:19456--9.

Vitousek, P. M., H. a Mooney, J. Lubchenco, and J. M. Melillo. 1997.
Human Domination of Earth' s Ecosystems. Science 277:494--499.

Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.

Wang, S., and M. Loreau. 2016. Biodiversity and ecosystem stability
across scales in metacommunities. Ecology Letters.

Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
in a fluctuating environment: the insurance hypothesis. Proceedings of
the National Academy of Sciences of the United States of America
96:1463--1468.

Yenni, G., P. B. Adler, and S. K. {Morgan Ernest}. 2012. Strong
self-limitation promotes the persistence of rare species. Ecology
93:456--461.

Zachmann, L., C. Moffet, and P. Adler. 2010. Mapped quadrats in
sagebrush steppe: long-term data for analyzing demographic rates and
plant--plant interactions. Ecology 91:3427.

\newpage{}

\subsection{Tables}\label{tables}

\singlespacing

Table 1: Site descriptions and focal species. \footnotesize

\begin{longtable}[c]{@{}lllll@{}}
\toprule
Site Name & Biome & Location (Lat, Lon) & Obs. Years &
Species\tabularnewline
\midrule
\endhead
New Mexico & Chihuahuan Desert & 32.62° N, 106.67° W & 1915-1950 &
\emph{Bouteloua eriopoda}\tabularnewline
& & & & \emph{Sporobolus flexuosus}\tabularnewline
Arizona & Sonoran Desert & 31°50' N, 110°53' W & 1915-1933 &
\emph{Bouteloua eriopoda}\tabularnewline
& & & & \emph{Bouteloua rothrockii}\tabularnewline
Kansas & Southern mixed prairie & 38.8° N, 99.3° W & 1932-1972 &
\emph{Bouteloua curtipendula}\tabularnewline
& & & & \emph{Bouteloua hirsuta}\tabularnewline
& & & & \emph{Schizachyrium scoparium}\tabularnewline
Montana & Northern mixed prairie & 46°19' N, 105°48' W & 1926-1957 &
\emph{Bouteloua gracilis}\tabularnewline
& & & & \emph{Hesperostipa comata}\tabularnewline
& & & & \emph{Pascopyrum smithii}\tabularnewline
& & & & \emph{Poa secunda}\tabularnewline
Idaho & Sagebrush steppe & 44.2° N, 112.1° W & 1926-1957 &
\emph{Artemisia tripartita}\tabularnewline
& & & & \emph{Pseudoroegneria spicata}\tabularnewline
& & & & \emph{Hesperostipa comata}\tabularnewline
& & & & \emph{Poa secunda}\tabularnewline
\bottomrule
\end{longtable}

\normalsize

\pagebreak{}

\begin{table}[ht]
\centering
\caption{Observed synchrony among species' per capita growth rates ($\phi_{R}$), observed synchrony among species' percent cover ($\phi_{C}$), the coefficient of variation of total community cover, and species richness for each community. Species richness values reflect the number of species analyzed from the community, not the actual richness.} 
{\normalsize
\begin{tabular}{lrrrr}
  \hline
Site & $\phi_{R}$ & $\phi_{C}$ & CV of Total Cover & Species richness \\ 
  \hline
New Mexico & 0.86 & 0.92 & 0.51 &   2 \\ 
  Arizona & 0.89 & 0.80 & 0.47 &   2 \\ 
  Kansas & 0.54 & 0.15 & 0.30 &   3 \\ 
  Montana & 0.53 & 0.54 & 0.52 &   4 \\ 
  Idaho & 0.36 & 0.18 & 0.19 &   4 \\ 
   \hline
\end{tabular}
}
\end{table}

\pagebreak{}

\subsection{Figures}\label{figures}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure1.png}
  \caption{Community-wide species synchrony of per capita growth rates from model simulation experiments. Synchrony of species' growth rates for each study area are from simulation experiments with demographic stochasticity, environmental stochasticity, and interspecific interactions present (``All Drivers''), demographic stochasticity removed (``No D.S.''), environmental stochasticity removed (``No E.S.''), interspecific interactions removed (``No Comp.''), interspecific interactions and demographic stochasticity removed (``No Comp. + No D.S.''), and interspecific interactions and environmental stochasticity removed (``No Comp. + No E.S.''). Abbreviations within the bars for the New Mexico site indicate whether the IBM or IPM was used for a particular simulation. \textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.} All IBM simulations shown in this figure were run on a 25 $\text{m}^2$ virtual landscape. Points show observed and predicted synchrony aligned with the model simulation that corresponds with each observation or analytical predicion.}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/formatted_figures/formatted_figure2.png}
  \caption{Synchrony of species' growth rates for each study area from IBM simulations across different landscape sizes when only demographic stochastcity is present (``D.S. Only'') and when environmental stochasticity is also present removed (``D.S. + E.S.''). The horizontal lines show the analytical predictions $\mathcal{M}_D$ (dashed line) and $\mathcal{M}_E$ (dotted line). The strength of demographic stochasticity decreases as landscape size increases because population sizes also increase. Theoretically, ``D.S. Only'' simulations should remain constant across landscape size, whereas ``D.S. + E.S.'' simulations should shift from the $\mathcal{M}_D$ prediction to the $\mathcal{M}_E$ prediction as landscape size, and thus population size, increases, but only if demographic stochasticity it strong enough to counteract environmental forcing. \textcolor{blue}{Error bars represent the 2.5\% and 97.5\% quantiles from model simulations.}}
\end{figure}

\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/formatted_figures/formatted_figure3.png}
  \caption{Synchrony of species per capita growth rates when species are growing in isolation (IPM without species interactions) versus synchrony in polycultures (IPM with species interactions). We used the same sequence of random year effects for both simulations (with and without species interactions) to mimic biodiversity-ecosystem functioning experiments. The dashed line is the line of equality. Simulation results in this figure are analagous to ``No Comp. + No D.S.'' (species in isolation) and ``No D.S.'' (species in competition) in Fig. 1, but here we control the time series of random year effects.}
\end{figure}

\end{document}